---
title: "CompareConversionEfficiency"
author: "Shelly Trigg"
date: "4/24/2020"
output: html_document
---

load library
```{r}
library(gtools)
library(dplyr)
library(tidyr)
library(broom)
library(purrr)
library(ggpubr)
```

read in data
```{r}
#lambda data (intermediate file)
df <- read.csv("OLD_Output/lamda_alignments_descriptive_stats.csv", stringsAsFactors = FALSE)

#Pact/Mcap data
df_pm <- read.csv("OLD_Output/Pact_Mcap_alignments_descriptive_stats.csv", stringsAsFactors = FALSE)

#meta data
meta_data <- read.csv("metadata/30-244831635_sample_submission_form.csv", skip = 1, header = TRUE, quote = "", row.names = NULL)
```

create column for total chh and total chg counts
```{r}
#first need to make column for total CHH counts
df_pm$tot_chh <- df_pm$meth_chh + df_pm$unmeth_chh

#first need to make column for total CHG counts
df_pm$tot_chg <- df_pm$meth_chg + df_pm$unmeth_chg

```

create columns for combined unmeth CHG and unmeth CHH, and for combined total CHH and CHG counts
```{r}
#combined totals
df_pm$tot_chh_chg <- df_pm$tot_chh + df_pm$tot_chg

#combined unmeth
df_pm$unmeth_chh_chg <- df_pm$unmeth_chh + df_pm$unmeth_chg

```

estimate conversion efficiency 
```{r}
#now calculate estimated conversion efficiency from %unmeth CHH
df_pm$estConvEffCHG_CHH <- (df_pm$unmeth_chh_chg/df_pm$tot_chh_chg)*100
```


merge dfs
```{r}
df_l <- merge(df[,c("Sample", "conversion_efficiency")], df_pm[,c("Sample","estConvEffCHG_CHH")], by = "Sample")

```

rename columns
```{r}
colnames(df_l) <- c("Sample", "conversion_efficiency_lambda", "est_conversion_efficiency_umCHG_umCHH")
```

combine with meta data
```{r}
#subset meta data for sample name, library prep, and species
meta_data <- meta_data[,c("Sample.Name.", "Species.Strain.", "Library.Prep.Kit.Used.")]

#re-name columns
colnames(meta_data) <- c("Sample", "Species", "Method")

#simplify method column
meta_data$Method <- gsub("MBD.*", "MBDBS", meta_data$Method)
meta_data$Method <- gsub(".* RRBS.*", "RRBS", meta_data$Method)
meta_data$Method <- gsub("Pico.*", "WGBS", meta_data$Method)

#merge meta data with conversion efficiency data

df_l <- merge(meta_data, df_l, by = "Sample")

#merge descriptive stats with conversion efficiency data
df_l <- merge(df_l, within(df,rm(conversion_efficiency)), by = "Sample")

#order table by sample

df_l <- df_l[mixedorder(df_l$Sample),]

```

write table
```{r}
write.csv(df_l, "OLD_Output/lamda_vs_estimated_conversion_efficiency.csv", row.names = FALSE, quote = FALSE)
```

convert to long format for ANOVA
```{r}
df_l_long <- gather(df_l, "convr_method", "perc_conv", 4:5)

#simplify convr method column
df_l_long$convr_method <- gsub(".*lambda", "lambda", df_l_long$convr_method)

df_l_long$convr_method <- gsub("est.*", "estimated", df_l_long$convr_method)
```

run anova on lambda conversion efficiency ~ BS method
```{r}

#run anova on lambda conversion efficiency ~ BS method for Pact samples and for Mcap samples
#The following code produces a list of anova model outputs for each species
aov_convr_lam <- df_l_long[which(df_l_long$convr_method=="lambda"),] %>% group_by(Species) %>% do(aov_models = aov(perc_conv ~ Method, data =.))

#make a tidy table with p.values
aov_convr_lam_tidy <- aov_convr_lam %>% ungroup %>% 
    pull(aov_models) %>% 
    map_dfr(tidy, .id = "grp")
#make a tidy table with glance stats
aov_convr_lam_glance <- aov_convr_lam %>% ungroup %>% 
    pull(aov_models) %>% 
    map_dfr(glance, .id = "grp")
#spread resid. data 
aov_convr_lam_tidy_w <- data.frame(pivot_wider(aov_convr_lam_tidy, names_from = term, values_from=c(df, sumsq, meansq, statistic, p.value)))

#remove columns with all NAs
aov_convr_lam_tidy_w <-aov_convr_lam_tidy_w[colSums(!is.na(aov_convr_lam_tidy_w)) > 0]

#merge all stats into one table
aov_convr_lam_stats <- merge(aov_convr_lam_glance,aov_convr_lam_tidy_w,by = "grp")

#add group names back in
aov_convr_lam_stats <- cbind(aov_convr_lam[,1], aov_convr_lam_stats[,-1])

#transpose df
aov_convr_lam_stats_t <- setNames(data.frame(t(aov_convr_lam_stats[,-1])), aov_convr_lam_stats[,1])

#add stats column
aov_convr_lam_stats_t$statistic <- as.character(rownames(aov_convr_lam_stats_t))

#remove rownames
rownames(aov_convr_lam_stats_t) <- NULL

#add model column
aov_convr_lam_stats_t$model <- "lambda conversion efficiency ~ library prep method"
```

run anova on estimated conversion efficiency ~ BS method
```{r}

#run anova on estimated conversion efficiency ~ BS method for Pact samples and for Mcap samples
#The following code produces a list of anova model outputs for each species
aov_convr_est <- df_l_long[which(df_l_long$convr_method=="estimated"),] %>% group_by(Species) %>% do(aov_models = aov(perc_conv ~ Method, data =.))

#make a tidy table with p.values
aov_convr_est_tidy <- aov_convr_est %>% ungroup %>% 
    pull(aov_models) %>% 
    map_dfr(tidy, .id = "grp")
#make a tidy table with glance stats
aov_convr_est_glance <- aov_convr_est %>% ungroup %>% 
    pull(aov_models) %>% 
    map_dfr(glance, .id = "grp")
#spread resid. data 
aov_convr_est_tidy_w <- data.frame(pivot_wider(aov_convr_est_tidy, names_from = term, values_from=c(df, sumsq, meansq, statistic, p.value)))

#remove columns with all NAs
aov_convr_est_tidy_w <-aov_convr_est_tidy_w[colSums(!is.na(aov_convr_est_tidy_w)) > 0]

#merge all stats into one table
aov_convr_est_stats <- merge(aov_convr_est_glance,aov_convr_est_tidy_w,by = "grp")

#add group names back in
aov_convr_est_stats <- cbind(aov_convr_est[,1], aov_convr_est_stats[,-1])

#transpose df
aov_convr_est_stats_t <- setNames(data.frame(t(aov_convr_est_stats[,-1])), aov_convr_est_stats[,1])

#stats column 
aov_convr_est_stats_t$statistic <- as.character(rownames(aov_convr_est_stats_t))

#remove rownames
rownames(aov_convr_est_stats_t) <- NULL

#add model column
aov_convr_est_stats_t$model <- " estimiated conversion efficiency ~ library prep method"



```






compare lambda to estimated conversion efficiency by species and method
```{r}
t_convr_cmpr <- df_l_long %>% group_by(Species, Method) %>% do(t_models = t.test(perc_conv ~ convr_method, data =.))

#make a tidy table with p.values
t_convr_cmpr_stats <- t_convr_cmpr %>% ungroup %>% 
    pull(t_models) %>% 
    map_dfr(tidy, .id = "grp")
#make a tidy table with glance stats
#aov_convr_cmpr_glance <- aov_convr_cmpr %>% ungroup %>% 
#    pull(aov_models) %>% 
#    map_dfr(glance, .id = "grp")
#spread resid. data 
#aov_convr_cmpr_tidy_w <- data.frame(pivot_wider(aov_convr_cmpr_tidy, names_from = term, values_from=c(df, sumsq, meansq, statistic, p.value)))

#remove columns with all NAs
#aov_convr_cmpr_tidy_w <-aov_convr_cmpr_tidy_w[colSums(!is.na(aov_convr_cmpr_tidy_w)) > 0]

#merge all stats into one table
#aov_convr_cmpr_stats <- merge(aov_convr_cmpr_glance,aov_convr_cmpr_tidy_w,by = "grp")

#add group names back in
t_convr_cmpr_stats <- cbind(t_convr_cmpr[,1:2], t_convr_cmpr_stats[,-1])

#add model column
t_convr_cmpr_stats$model <- "estimated ~ lambda"

#add lib. prep method to model
t_convr_cmpr_stats$model <- paste(t_convr_cmpr_stats$Method, t_convr_cmpr_stats$model, t_convr_cmpr_stats$alternative, t_convr_cmpr_stats$method)

#remove lib method column
t_convr_cmpr_stats$Method <- NULL
t_convr_cmpr_stats$method <- NULL
t_convr_cmpr_stats$alternative <- NULL

#transpose df
#t_convr_cmpr_stats_t <- setNames(data.frame(t(t_convr_cmpr_stats[,-c(1:2)])),c(t_convr_cmpr_stats[,1]))

#format to long
t_convr_cmpr_stats_t <- tidyr::spread(tidyr::gather( t_convr_cmpr_stats, "statistic", "value", 2:9), "Species", "value")

```

write out ANOVA stats table
```{r}
#combine summary tables
aov_convr_stats_all <- rbind(aov_convr_lam_stats_t,aov_convr_est_stats_t,t_convr_cmpr_stats_t)

write.csv(aov_convr_stats_all, "OLD_Output/conversion_efficiency_stats.csv", quote = FALSE)
```


generate plots to match ANOVAs
```{r}
#plot lambda conversion efficiency for each method facetted by species
jpeg("OLD_Output/BS_conversion_efficiency.jpg", width = 6, height = 3, units = "in", res = 300)
ggplot(df_l_long) + geom_boxplot(aes(y = perc_conv, x = Method, color = Method, fill = convr_method),outlier.shape = NA) + facet_wrap(~Species) + theme_bw() + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),strip.background =element_rect(fill="white"),strip.text = element_text(face = "italic")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + ylab("conversion efficiency (%)") + scale_fill_grey() + labs(color = "library prep method",fill = "conversion efficiency\ncalculation method")
dev.off()
```

