---
title: "Untitled"
author: "Shelly Trigg"
date: "5/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

load libraries
```{r}
library(data.table)
library(ggplot2)
library(dplyr)
library(drc)
library(ggpubr)

```


read in data
```{r}
#data for percent genome coverage
#these were generated by https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/20200416_CollectWgsMetrics.sh 

Meth1 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth1_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth2 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth2_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth3 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth3_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth4 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth4_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth5 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth5_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth6 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth6_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth7 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth7_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth8 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth8_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth9 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Pact/Meth9_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

#Mcap
Meth10 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth10_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth11 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth11_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth12 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth12_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth13 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth13_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth14 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth14_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth15 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth15_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Meth16 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth16_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth17 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth17_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Meth18 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200416/Mcap/Meth18_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")



# data for saturation estimates
#these data were generated by https://gannet.fish.washington.edu/metacarcinus/mox_jobs/20200513_SubsamplePicardPact.sh and https://gannet.fish.washington.edu/metacarcinus/mox_jobs/20200527_SubsamplePicardMcap.sh

Pact_WGBS_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_WGBS_50M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_WGBS_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_WGBS_100M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_WGBS_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_WGBS_150M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_WGBS_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_WGBS_200M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Pact_RRBS_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_RRBS_50M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Pact_RRBS_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_RRBS_100M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Pact_RRBS_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_RRBS_150M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Pact_RRBS_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_RRBS_200M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Pact_MBD_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_MBD_50M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_MBD_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_MBD_100M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_MBD_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_MBD_150M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Pact_MBD_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200513/Pact_MBD_200M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")


#Mcap
Mcap_WGBS_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_WGBS_50M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_WGBS_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_WGBS_100M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_WGBS_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_WGBS_150M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_WGBS_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_WGBS_200M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

Mcap_RRBS_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_RRBS_50M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Mcap_RRBS_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_RRBS_100M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Mcap_RRBS_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_RRBS_150M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")
Mcap_RRBS_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_RRBS_200M_R1_001_val_1_bismark_bt2_pe.collect_wgs_metrics.txt")

Mcap_MBD_50 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_MBD_50M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_MBD_100 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_MBD_100M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_MBD_150 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_MBD_150M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")
Mcap_MBD_200 <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_MBD_200M_R1_001_val_1_bismark_bt2_pe.deduplicated.collect_wgs_metrics.txt")

#CpG data for samples and for pooled samples
#These data were generated by https://github.com/hputnam/Meth_Compare/blob/master/scripts/Pact_CpG_coverageXdepth.ipynb and https://github.com/hputnam/Meth_Compare/blob/master/scripts/Mcap_CpG_coverageXdepth.ipynb . The CSV files were created from copying and pasting numbers from jupyter output.

Pact_Sample_CpG <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200514/Pact_CpG_coverage.csv")

Pact_pool_Sample_CpG <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200514/Pact_CpG_est_coverage.csv")

#Mcap 

Mcap_Sample_CpG <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200527/Mcap_CpG_coverage.csv")

Mcap_pool_Sample_CpG <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200725/Mcap_cpg_est_coverage.csv")

```

Add meta data 
```{r}
#data for percent genome coverage
# add method column
Meth1$method <- "WGBS"
Meth2$method <- "WGBS"
Meth3$method <- "WGBS"
Meth4$method <- "RRBS"
Meth5$method <- "RRBS"
Meth6$method <- "RRBS"
Meth7$method <- "MBD"
Meth8$method <- "MBD"
Meth9$method <- "MBD"

Meth10$method <- "WGBS"
Meth11$method <- "WGBS"
Meth12$method <- "WGBS"
Meth13$method <- "RRBS"
Meth14$method <- "RRBS"
Meth15$method <- "RRBS"
Meth16$method <- "MBD"
Meth17$method <- "MBD"
Meth18$method <- "MBD"

# data for saturation estimates
#add group and depth info to each dataframe

#pact 

Pact_WGBS_50$method <- "WGBS"
Pact_WGBS_50$depth <- "50"

Pact_WGBS_100$method <- "WGBS"
Pact_WGBS_100$depth <- "100"

Pact_WGBS_150$method <- "WGBS"
Pact_WGBS_150$depth <- "150"

Pact_WGBS_200$method <- "WGBS"
Pact_WGBS_200$depth <- "200"

Pact_RRBS_50$method <- "RRBS"
Pact_RRBS_50$depth <- "50"

Pact_RRBS_100$method <- "RRBS"
Pact_RRBS_100$depth <- "100"

Pact_RRBS_150$method <- "RRBS"
Pact_RRBS_150$depth <- "150"

Pact_RRBS_200$method <- "RRBS"
Pact_RRBS_200$depth <- "200"

Pact_MBD_50$method <- "MBD"
Pact_MBD_50$depth <- "50"

Pact_MBD_100$method <- "MBD"
Pact_MBD_100$depth <- "100"

Pact_MBD_150$method <- "MBD"
Pact_MBD_150$depth <- "150"

Pact_MBD_200$method <- "MBD"
Pact_MBD_200$depth <- "200"

#Mcap

Mcap_WGBS_50$method <- "WGBS"
Mcap_WGBS_50$depth <- "50"

Mcap_WGBS_100$method <- "WGBS"
Mcap_WGBS_100$depth <- "100"

Mcap_WGBS_150$method <- "WGBS"
Mcap_WGBS_150$depth <- "150"

Mcap_WGBS_200$method <- "WGBS"
Mcap_WGBS_200$depth <- "200"

Mcap_RRBS_50$method <- "RRBS"
Mcap_RRBS_50$depth <- "50"

Mcap_RRBS_100$method <- "RRBS"
Mcap_RRBS_100$depth <- "100"

Mcap_RRBS_150$method <- "RRBS"
Mcap_RRBS_150$depth <- "150"

Mcap_RRBS_200$method <- "RRBS"
Mcap_RRBS_200$depth <- "200"

Mcap_MBD_50$method <- "MBD"
Mcap_MBD_50$depth <- "50"

Mcap_MBD_100$method <- "MBD"
Mcap_MBD_100$depth <- "100"

Mcap_MBD_150$method <- "MBD"
Mcap_MBD_150$depth <- "150"

Mcap_MBD_200$method <- "MBD"
Mcap_MBD_200$depth <- "200"


# CpG data for samples
# add method column

#Pact

for(i in 1:nrow(Pact_Sample_CpG)){
  if(Pact_Sample_CpG$Sample[i] == "Meth_1" |Pact_Sample_CpG$Sample[i] == "Meth_2" | Pact_Sample_CpG$Sample[i] == "Meth_3"){
    Pact_Sample_CpG$method[i] <- "WGBS"
  }
  if(Pact_Sample_CpG$Sample[i] == "Meth_4" |Pact_Sample_CpG$Sample[i] == "Meth_5" | Pact_Sample_CpG$Sample[i] == "Meth_6"){
    Pact_Sample_CpG$method[i] <- "RRBS"
  }
  if(Pact_Sample_CpG$Sample[i] == "Meth_7" |Pact_Sample_CpG$Sample[i] == "Meth_8" | Pact_Sample_CpG$Sample[i] == "Meth_9"){
    Pact_Sample_CpG$method[i] <- "MBD"
  }
}

#Mcap
for(i in 1:nrow(Mcap_Sample_CpG)){
  if(Mcap_Sample_CpG$Sample[i] == "Meth_10" |Mcap_Sample_CpG$Sample[i] == "Meth_11" | Mcap_Sample_CpG$Sample[i] == "Meth_12"){
    Mcap_Sample_CpG$method[i] <- "WGBS"
  }
  if(Mcap_Sample_CpG$Sample[i] == "Meth_13" |Mcap_Sample_CpG$Sample[i] == "Meth_14" | Mcap_Sample_CpG$Sample[i] == "Meth_15"){
    Mcap_Sample_CpG$method[i] <- "RRBS"
  }
  if(Mcap_Sample_CpG$Sample[i] == "Meth_16" |Mcap_Sample_CpG$Sample[i] == "Meth_17" | Mcap_Sample_CpG$Sample[i] == "Meth_18"){
    Mcap_Sample_CpG$method[i] <- "MBD"
  }
}
```

combine sample data frames
```{r}
#data for percent genome coverage
Pact_samples <- rbind(Meth1,Meth2, Meth3, Meth4, Meth5, Meth6, Meth7, Meth8, Meth9)

Mcap_samples <- rbind(Meth10,Meth11, Meth12, Meth13, Meth14, Meth15, Meth16, Meth17, Meth18)

# data for saturation estimates

Pact_all <- rbind(Pact_WGBS_50, Pact_WGBS_100, Pact_WGBS_150, Pact_WGBS_200, Pact_RRBS_50, Pact_RRBS_100, Pact_RRBS_150, Pact_RRBS_200, Pact_MBD_50, Pact_MBD_100, Pact_MBD_150, Pact_MBD_200)

Mcap_all <- rbind(Mcap_WGBS_50, Mcap_WGBS_100, Mcap_WGBS_150, Mcap_WGBS_200, Mcap_RRBS_50, Mcap_RRBS_100, Mcap_RRBS_150, Mcap_RRBS_200, Mcap_MBD_50, Mcap_MBD_100, Mcap_MBD_150, Mcap_MBD_200)

```

format for plotting
```{r}
#data for percent genome coverage
#reformat to long
Pact_samples_stacked <- tidyr::gather(Pact_samples, "coverage", "percent_genome", 13:26)
#remove  PCT_ from coverage column
Pact_samples_stacked$coverage <- gsub("PCT_","", Pact_samples_stacked$coverage)
#remove X from coverage column
Pact_samples_stacked$coverage <- gsub("X","", Pact_samples_stacked$coverage)
#make coverage a numeric class
Pact_samples_stacked$coverage <- as.numeric(Pact_samples_stacked$coverage)



# data for saturation estimates
#reformat to long
Pact_all_stacked <- tidyr::gather(Pact_all, "coverage", "percent_genome", 13:26)
#remove PCT_ from coverage column
Pact_all_stacked$coverage <- gsub("PCT_","", Pact_all_stacked$coverage)
#remove X from coverage column
Pact_all_stacked$coverage <- gsub("X","", Pact_all_stacked$coverage)
#make coverage a numeric class
Pact_all_stacked$coverage <- as.numeric(Pact_all_stacked$coverage)
#reorder depth column and change to class factor for plotting
Pact_all_stacked$depth <- factor(Pact_all_stacked$depth, levels = c("200", "150", "100", "50"))

#CpG sample data
#reformat to long
Pact_Sample_CpG_stacked <- tidyr::gather(Pact_Sample_CpG, "coverage", "percent_CpG", 2:15)
#remove X from coverage column
Pact_Sample_CpG_stacked$coverage <- gsub("x","", Pact_Sample_CpG_stacked$coverage)
#make coverage a numeric class
Pact_Sample_CpG_stacked$coverage <- as.numeric(Pact_Sample_CpG_stacked$coverage)
#calculate percent CpG by dividing number of CpG by total number CpG in genome (9155620)
Pact_Sample_CpG_stacked$percent_CpG <- (Pact_Sample_CpG_stacked$percent_CpG/9155620)
#convert MBD to MBDBS
Pact_Sample_CpG_stacked$method <- gsub("MBD", "MBDBS", Pact_Sample_CpG_stacked$method)

#CpGs for pooled sample data
#reformat to long
Pact_pool_Sample_CpG_stacked <-  tidyr::gather(Pact_pool_Sample_CpG, "coverage", "percent_CpG", 3:16)
#remove X from coverage column
Pact_pool_Sample_CpG_stacked$coverage <- gsub("x","", Pact_pool_Sample_CpG_stacked$coverage)
#make coverage a numeric class
Pact_pool_Sample_CpG_stacked$coverage <- as.numeric(Pact_pool_Sample_CpG_stacked$coverage)
#calculate percent CpG by dividing number of CpG by total number CpG in genome (9155620)
Pact_pool_Sample_CpG_stacked$percent_CpG <- (Pact_pool_Sample_CpG_stacked$percent_CpG/9155620)
#reorder depth column and change to class factor for plotting
Pact_pool_Sample_CpG_stacked$depth <- factor(Pact_pool_Sample_CpG_stacked$depth, levels = c("200", "150", "100", "50"))
#convert MBD to MBDBS
Pact_pool_Sample_CpG_stacked$method <- gsub("MBD", "MBDBS", Pact_pool_Sample_CpG_stacked$method)

#Mcap

#data for percent genome coverage
#reformat to long
Mcap_samples_stacked <- tidyr::gather(Mcap_samples, "coverage", "percent_genome", 13:26)
#remove PCT_ from coverage column
Mcap_samples_stacked$coverage <- gsub("PCT_","", Mcap_samples_stacked$coverage)
#remove X from coverage column
Mcap_samples_stacked$coverage <- gsub("X","", Mcap_samples_stacked$coverage)
#make coverage a numeric class
Mcap_samples_stacked$coverage <- as.numeric(Mcap_samples_stacked$coverage)



# data for saturation estimates
#reformat to long
Mcap_all_stacked <- tidyr::gather(Mcap_all, "coverage", "percent_genome", 13:26)
#remove PCT_ from coverage column
Mcap_all_stacked$coverage <- gsub("PCT_","", Mcap_all_stacked$coverage)
#remove X from coverage column
Mcap_all_stacked$coverage <- gsub("X","", Mcap_all_stacked$coverage)
#make coverage a numeric class
Mcap_all_stacked$coverage <- as.numeric(Mcap_all_stacked$coverage)
#reorder depth column and change to class factor for plotting
Mcap_all_stacked$depth <- factor(Mcap_all_stacked$depth, levels = c("200", "150", "100", "50"))

#CpG sample data
#reformat to long
Mcap_Sample_CpG_stacked <- tidyr::gather(Mcap_Sample_CpG, "coverage", "percent_CpG", 2:15)
#remove X from coverage column
Mcap_Sample_CpG_stacked$coverage <- gsub("x","", Mcap_Sample_CpG_stacked$coverage)
#make coverage a numeric class
Mcap_Sample_CpG_stacked$coverage <- as.numeric(Mcap_Sample_CpG_stacked$coverage)
#calculate percent CpG by dividing number of CpG by total number CpG in genome (9155620)
Mcap_Sample_CpG_stacked$percent_CpG <- (Mcap_Sample_CpG_stacked$percent_CpG/28684519)
#convert MBD to MBDBS
Mcap_Sample_CpG_stacked$method <- gsub("MBD", "MBDBS", Mcap_Sample_CpG_stacked$method)


#CpGs for pooled sample data
#reformat to long
Mcap_pool_Sample_CpG_stacked <-  tidyr::gather(Mcap_pool_Sample_CpG, "coverage", "percent_CpG", 3:11)
#remove X from coverage column
Mcap_pool_Sample_CpG_stacked$coverage <- gsub("x","", Mcap_pool_Sample_CpG_stacked$coverage)
#make coverage a numeric class
Mcap_pool_Sample_CpG_stacked$coverage <- as.numeric(Mcap_pool_Sample_CpG_stacked$coverage)
#calculate percent CpG by dividing number of CpG by total number CpG in genome (28684519)
Mcap_pool_Sample_CpG_stacked$percent_CpG <- (Mcap_pool_Sample_CpG_stacked$percent_CpG/28684519)
#remove M from depth column
Mcap_pool_Sample_CpG_stacked$depth <- gsub("M","",Mcap_pool_Sample_CpG_stacked$depth)

#reorder depth column and change to class factor for plotting
Mcap_pool_Sample_CpG_stacked$depth <- factor(Mcap_pool_Sample_CpG_stacked$depth, levels = c("200", "150", "100", "50"))
#convert MBD to MBDBS
Mcap_pool_Sample_CpG_stacked$method <- gsub("MBD", "MBDBS", Mcap_pool_Sample_CpG_stacked$method)
```

calculate average and sd for Genome and CpGs covered, subset to 50x coverage
```{r}
#Mcap genome
Mcap_samples_avg <- Mcap_samples_stacked %>% group_by(method, coverage) %>% summarise(mean = mean(percent_genome), sd = sd(percent_genome))

#Pact genome
Pact_samples_avg <- Pact_samples_stacked %>% group_by(method, coverage) %>% summarise(mean = mean(percent_genome), sd = sd(percent_genome))

#Mcap CpGs
Mcap_Sample_CpG_avg <- data.frame(Mcap_Sample_CpG_stacked[which(Mcap_Sample_CpG_stacked$coverage<=50),]) %>% group_by(method, coverage) %>% summarise(mean = mean(percent_CpG), sd = sd(percent_CpG))

#Pact CpGs
Pact_Sample_CpG_avg <- data.frame(Pact_Sample_CpG_stacked[which(Pact_Sample_CpG_stacked$coverage<=50),]) %>% group_by(method, coverage) %>% summarise(mean = mean(percent_CpG), sd = sd(percent_CpG))

```

plot coverage
```{r}
# fraction of genome covered x method with loess smoothing
#jpeg("Output/Pact_GenomecovXmethod_sample_avg.jpg", width = 4.5, height = 3, units = "in", res = 300)
#ggplot(Pact_samples_stacked[which(Pact_samples_stacked$coverage <= 50),], aes(x = coverage, y = percent_genome, color = method)) + geom_smooth(aes(fill = method)) + scale_x_continuous(breaks = seq(0,100,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of genome") + theme_bw() + theme(text = element_text(size=16))
#dev.off()

# mean fraction of genome covered x method with sd
# aa <- ggplot(Mcap_samples_avg, aes(x = coverage, y = mean, color = method)) + geom_line(size = 0.7) + geom_ribbon(aes(ymin=mean-sd,ymax=mean + sd,fill=method),color="grey70",alpha=0.3, size = 0.1) + scale_x_continuous(breaks = seq(0,50,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of genome") + theme_bw() + theme(text = element_text(size=16))


#fraction of CpGs covered x method with loess smoothing
# jpeg("Output/Pact_CpGcovXmethod_sample_avg.jpg", width = 4.5, height = 3, units = "in", res = 300)
# ggplot(Pact_Sample_CpG_stacked[which(Pact_Sample_CpG_stacked$coverage <= 50),], aes(x = coverage, y = percent_CpG, color = method)) + geom_smooth(aes(fill = method)) + scale_x_continuous(breaks = seq(0,100,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=16))
# dev.off()

#mean fraction of CpGs covered x method with sd
cc <- ggplot(Pact_Sample_CpG_avg, aes(x = coverage, y = mean, color = method)) + geom_line(size = 1, alpha = 1) + geom_ribbon(aes(ymin=mean-sd,ymax=mean + sd,fill=method),color="grey70",alpha=0.5, size = 0.1) + scale_x_continuous(breaks = seq(0,50,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=16))

#inset (zoomed in on 5x and 10x CpG coverage)
ccc <- ggplot(Pact_Sample_CpG_avg[which(Pact_Sample_CpG_avg$coverage==5|Pact_Sample_CpG_avg$coverage==10),], aes(x = coverage, y = mean, color = method)) + geom_line(size = 1, alpha = 1) + geom_ribbon(aes(ymin=mean-sd,ymax=mean + sd,fill=method),color="grey70",alpha=0.5, size = 0.1) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=12),legend.position = "none") + scale_x_continuous(breaks = c(5,10))

#plot plus inset (zoomed in on 5x and 10x CpG coverage)
c <- cc + annotation_custom(ggplotGrob(ccc), xmin = 15, xmax = 50, ymin = 0.25, ymax = 0.88)

# fraction of genome covered by pooled samples
#jpeg("Output/Pact_GenomecovXmethod_pooled_subsamples.jpg", width = 9, height = 3, units = "in", res = 300)
#ggplot(Pact_all_stacked[which(Pact_all_stacked$coverage<=50),], aes(x = coverage, y= percent_genome, group = paste(method, depth, sep = "_"),color = depth)) + geom_line(size = 1, alpha = 0.8) + facet_wrap(~method)+  scale_x_continuous(breaks = seq(0,100,10))+ xlab("coverage (X)") + ylab("fraction of genome") + labs(color = "depth (M reads)") + scale_color_grey() + theme_bw() + theme(text = element_text(size=16)) + theme(strip.background =element_rect(fill="white"))
#dev.off()

# fraction of CpGs covered by pooled samples

#jpeg("Output/Pact_CpGcovXmethod_pooled_subsamples.jpg", width = 9, height = 3, units = "in", res = 300)
d <- ggplot(Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage<=50),], aes(x = coverage, y= percent_CpG, group = paste(method, depth, sep = "_"),color = depth)) + geom_line(size = 1, alpha = 0.8) + facet_wrap(~method)+  scale_x_continuous(breaks = seq(0,100,10))+ xlab("coverage (X)") + ylab("fraction of CpGs") + labs(color = "depth (M reads)") + scale_color_grey() + theme_bw() + theme(text = element_text(size=16)) + theme(strip.background =element_rect(fill="white"))
#dev.off()

#Mcap

# fraction of genome covered x method
#jpeg("Output/Mcap_GenomecovXmethod_sample_avg.jpg", width = 4.5, height = 3, units = "in", res = 300)
#ggplot(Mcap_samples_stacked[which(Mcap_samples_stacked$coverage <= 50),], aes(x = coverage, y = percent_genome, color = method)) + geom_smooth(aes(fill = method)) + scale_x_continuous(breaks = seq(0,100,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of genome") + theme_bw() + theme(text = element_text(size=16))
#dev.off()

#fraction of CpGs covered x method loess smoothing
#jpeg("Output/Mcap_CpGcovXmethod_sample_avg.jpg", width = 4.5, height = 3, units = "in", res = 300)
#ggplot(Mcap_Sample_CpG_stacked[which(Mcap_Sample_CpG_stacked$coverage <= 50),], aes(x = coverage, y = percent_CpG, color = method)) + geom_smooth(aes(fill = method)) + scale_x_continuous(breaks = seq(0,100,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=16))
#dev.off()

#mean fraction of CpGs covered x method with sd
aa <- ggplot(Mcap_Sample_CpG_avg, aes(x = coverage, y = mean, color = method)) + geom_line(size = 1, alpha = 1) + geom_ribbon(aes(ymin=mean-sd,ymax=mean + sd,fill=method),color="grey70",alpha=0.5, size = 0.1) + scale_x_continuous(breaks = seq(0,50,10)) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=16))



aaa <- ggplot(Mcap_Sample_CpG_avg[which(Mcap_Sample_CpG_avg$coverage==5|Mcap_Sample_CpG_avg$coverage==10),], aes(x = coverage, y = mean, color = method)) + geom_line(size = 1, alpha = 1) + geom_ribbon(aes(ymin=mean-sd,ymax=mean + sd,fill=method),color="grey70",alpha=0.5, size = 0.1) + scale_fill_manual(values = c("#E6550D", "#756BB1","#31A354")) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + xlab("coverage (X)") + ylab("fraction of CpGs") + theme_bw() + theme(text = element_text(size=12),legend.position = "none") + scale_x_continuous(breaks = c(5,10))

a <- aa + annotation_custom(ggplotGrob(aaa), xmin = 10, xmax = 50, ymin = 0.15, ymax = 0.7)

# fraction of genome covered by pooled samples
#jpeg("Output/Mcap_GenomecovXmethod_pooled_subsamples.jpg", width = 9, height = 3, units = "in", res = 300)
#ggplot(Mcap_all_stacked[which(Mcap_all_stacked$coverage<=50),], aes(x = coverage, y= percent_genome, group = paste(method, depth, sep = "_"),color = depth)) + geom_line(size = 1, alpha = 0.8) + facet_wrap(~method)+  scale_x_continuous(breaks = seq(0,100,10))+ xlab("coverage (X)") + ylab("fraction of genome") + labs(color = "depth (M reads)") + scale_color_grey() + theme_bw() + theme(text = element_text(size=16)) + theme(strip.background =element_rect(fill="white"))
#dev.off()

# fraction of CpGs covered by pooled samples

#jpeg("Output/Mcap_CpGcovXmethod_pooled_subsamples.jpg", width = 9, height = 3, units = "in", res = 300)
b <- ggplot(Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage<=50),], aes(x = coverage, y= percent_CpG, group = paste(method, depth, sep = "_"),color = depth)) + geom_line(size = 1, alpha = 0.8) + facet_wrap(~method)+  scale_x_continuous(breaks = seq(0,100,10))+ xlab("coverage (X)") + ylab("fraction of CpGs") + labs(color = "depth (M reads)") + scale_color_grey() + theme_bw() + theme(text = element_text(size=16)) + theme(strip.background =element_rect(fill="white"))
#dev.off()

jpeg("OLD_Output/Mcap_Pact_gnm-wide_CpGcov.jpg", width = 15, height = 7, units = "in", res = 300)
ggarrange(a,b,c,d, labels = c("A", "B", "C", "D"),font.label = list(size = 24),nrow = 2,ncol = 2,align = "hv",widths = c(1,2))
dev.off()

```








model coverage needed to saturate CpGs covered
```{r}
#create object that is a vector of the methods
methods <- unique(Mcap_pool_Sample_CpG_stacked$method)
#create empty df
predicted_CpG_cov_all_methods <- data.frame()
#create an empty data frame to save the model info in
model_df <- data.frame()
#loop through data 
for (i in 1:length(methods)){
  # use the two parameter MM model (MM.2)
  m1 <- drm(percent_CpG ~ as.integer(as.character(depth)), data =Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),], fct = MM.2())
  #create vmax object from model
  Vmax <- data.frame(tibble(coef(m1)))[1,]
  #create K object from model
  Km <- data.frame(tibble(coef(m1)))[2,]
  #create model df to save model for each iteration
  mm <- data.frame(cbind(summary(m1)[[3]],rbind(summary(m1)[[13]], summary(m1)[[13]])))
  #create a Vmax and Km SE high and low
  Vmax_SE_low <- Vmax - mm$Std..Error[1]
  Vmax_SE_high <- Vmax + mm$Std..Error[1]
  Km_SE_low <- Km - mm$Std..Error[2]
  Km_SE_high <- Km + mm$Std..Error[2]
  #add method column
  mm$method <- rep(methods[i],2)
  #combine model info with model info master df
  model_df <- rbind(model_df, mm)
  #create empty df
  predicted_CpG_cov <- data.frame()
  #create a series of numbers 50-3000 by 50s that will be the estimated depth
  prediction <- seq(50,4000, by = 50)
  #loop through the depths 
  for (j in 1:length(prediction)){
    #calculate the estimated percent CpGs covered (using MM model coefficients)
    perc_CpG <- (Vmax*prediction[j]/(Km+prediction[j]))
    # calculate perc_CpG for SE high and low
    perc_CpG_SE_high <- (Vmax_SE_high*prediction[j]/(Km_SE_high+prediction[j]))
    perc_CpG_SE_low <- (Vmax_SE_low*prediction[j]/(Km_SE_low+prediction[j]))
    #save the depth as an object
    depth <- prediction[j]
    #save the percent CpGs covered and the depth as a data frame
    pred_perc_CpG <- data.frame(perc_CpG,perc_CpG_SE_low,perc_CpG_SE_high,depth)
    #combine data frame for each depth in a master data frame
    predicted_CpG_cov <- rbind(predicted_CpG_cov,pred_perc_CpG)
  }
  #create column for denoting series
  #predicted_CpG_cov <- tidyr::gather(predicted_CpG_cov,"subseries", "perc_CpG", 1:3)
  #predicted_CpG_cov$subseries <- gsub("perc_CpG", "estimated", predicted_CpG_cov$subseries)
  predicted_CpG_cov$series <- "estimated"
  #create an object for the actual CpGs covered at known depths
  actual_CpG_cov <- data.frame(perc_CpG = Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),c("percent_CpG")],depth = as.integer(as.character(Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),c("depth")])),series = "observed")
  #actual_CpG_cov$subseries <- "observed"
  actual_CpG_cov$perc_CpG_SE_low <- NA
  actual_CpG_cov$perc_CpG_SE_high <- NA
  pred_act_cov <- rbind(predicted_CpG_cov, actual_CpG_cov) 
  pred_act_cov$method <- methods[i]
  predicted_CpG_cov_all_methods <- rbind(predicted_CpG_cov_all_methods,pred_act_cov)
}

#threshold Vmax to 1
predicted_CpG_cov_all_methods$perc_CpG_SE_high[predicted_CpG_cov_all_methods$perc_CpG_SE_high > 1] <- 1

model_df$coefficient <- rownames(model_df)
model_df$coefficient <- gsub("d.*","Vmax", model_df$coefficient)
model_df$coefficient <- gsub("e.*", "Km", model_df$coefficient)

mcap_model_df <- model_df %>% tidyr::pivot_longer(names_to = "statistic", values_to = "value",1:6)%>% tidyr::pivot_wider(names_from = coefficient, values_from = value)

mcap_model_df$species <- "M.capitata"

```

plot saturation curves per method
```{r}
#a <- ggplot(predicted_CpG_cov_all_methods,aes(x = depth, y = perc_CpG, color = method, alpha = series)) + geom_point() + geom_pointrange(aes(ymin=perc_CpG_SE_low, ymax=perc_CpG_SE_high)) + scale_x_continuous(breaks = seq(0,4000,100))+ scale_y_continuous(labels = as.character(seq(0,1.0,0.2)),breaks = seq(0,1,0.2), limits = c(0,1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.background =element_rect(fill="white"), plot.title =  element_text(face = "italic")) + ylab("fraction of genome-wide\nCpGs covered") + xlab("depth (M reads)") + labs(color = "method" ) + ggtitle("M. capitata") + scale_color_manual(values =c("#E6550D","#756BB1","#31A354")) + scale_alpha_manual(values=c(0.3, 1))


a <- ggplot(predicted_CpG_cov_all_methods,aes(x = depth, y = perc_CpG, color = method, alpha = series)) + geom_point(size = 0.5) + geom_line() + geom_ribbon(aes(ymin=perc_CpG_SE_low,ymax=perc_CpG_SE_high,fill=method),color="grey70",alpha=0.2, size = 0.1) + scale_x_continuous(breaks = seq(0,4000,100))+ scale_y_continuous(labels = as.character(seq(0,1.0,0.2)),breaks = seq(0,1,0.2)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.background =element_rect(fill="white"), plot.title =  element_text(face = "italic")) + ylab("fraction of genome-wide\nCpGs covered") + xlab("depth (M reads)") + labs(color = "method" ) + ggtitle("M. capitata") + scale_color_manual(values =c("#E6550D","#756BB1","#31A354")) + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354")) +  scale_alpha_manual(values=c(0.3, 1))
```

do the same for pact
model coverage needed to saturate CpGs covered

```{r}
#create object that is a vector of the methods
methods <- unique(Pact_pool_Sample_CpG_stacked$method)
#create empty df
predicted_CpG_cov_all_methods <- data.frame()
#create an empty data frame to save the model info in
model_df <- data.frame()
#loop through data 
for (i in 1:length(methods)){
  # use the two parameter MM model (MM.2)
  m1 <- drm(percent_CpG ~ as.integer(as.character(depth)), data =Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage==5 & Pact_pool_Sample_CpG_stacked$method == methods[i]),], fct = MM.2())
  #create vmax object from model
  Vmax <- data.frame(tibble(coef(m1)))[1,]
  #create K object from model
  Km <- data.frame(tibble(coef(m1)))[2,]
  #create model df to save model for each iteration
  mm <- data.frame(cbind(summary(m1)[[3]],rbind(summary(m1)[[13]], summary(m1)[[13]])))
  #create a Vmax and Km SE high and low
  Vmax_SE_low <- Vmax - mm$Std..Error[1]
  Vmax_SE_high <- Vmax + mm$Std..Error[1]
  Km_SE_low <- Km - mm$Std..Error[2]
  Km_SE_high <- Km + mm$Std..Error[2]
  #add method column
  mm$method <- rep(methods[i],2)
  #combine model info with model info master df
  model_df <- rbind(model_df, mm)
  #create empty df
  predicted_CpG_cov <- data.frame()
  #create a series of numbers 50-3000 by 50s that will be the estimated depth
  prediction <- seq(50,4000, by = 50)
  #loop through the depths 
  for (j in 1:length(prediction)){
    #calculate the estimated percent CpGs covered (using MM model coefficients)
    if(Vmax > 1){
      x <- Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage==5 & Pact_pool_Sample_CpG_stacked$method == methods[i]),"percent_CpG"]
      y <- which(abs(x-0.5)==min(abs(x-0.5)))
      Km <- as.numeric(as.character(Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage==5 & Pact_pool_Sample_CpG_stacked$method == methods[i] & Pact_pool_Sample_CpG_stacked$percent_CpG == x[[y]]),"depth"]))
    perc_CpG <- (1*prediction[j]/(Km+prediction[j]))
     # calculate perc_CpG for SE high and low
    perc_CpG_SE_high <- ((1+mm$Std..Error[1])*prediction[j]/(Km+prediction[j]))
    perc_CpG_SE_low <- ((1-mm$Std..Error[1])*prediction[j]/(Km_SE_low+prediction[j]))
    #save the depth as an object
    depth <- prediction[j]
    #save the percent CpGs covered and the depth as a data frame
    pred_perc_CpG <- data.frame(perc_CpG, perc_CpG_SE_low,perc_CpG_SE_high,depth)
    #combine data frame for each depth in a master data frame
    predicted_CpG_cov <- rbind(predicted_CpG_cov,pred_perc_CpG)
    }
    else{
    perc_CpG <- (Vmax*prediction[j]/(Km+prediction[j]))
     # calculate perc_CpG for SE high and low
    perc_CpG_SE_high <- (Vmax_SE_high*prediction[j]/(Km_SE_high+prediction[j]))
    perc_CpG_SE_low <- (Vmax_SE_low*prediction[j]/(Km_SE_low+prediction[j]))
    
    #save the depth as an object
    depth <- prediction[j]
    #save the percent CpGs covered and the depth as a data frame
    pred_perc_CpG <- data.frame(perc_CpG, perc_CpG_SE_low, perc_CpG_SE_high,depth)
    #combine data frame for each depth in a master data frame
    predicted_CpG_cov <- rbind(predicted_CpG_cov,pred_perc_CpG)
    }
  }
  predicted_CpG_cov$series <- "estimated"
  #create an object for the actual CpGs covered at known depths
  actual_CpG_cov <- data.frame(perc_CpG = Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage==5 & Pact_pool_Sample_CpG_stacked$method == methods[i]),c("percent_CpG")],depth = as.integer(as.character(Pact_pool_Sample_CpG_stacked[which(Pact_pool_Sample_CpG_stacked$coverage==5 & Pact_pool_Sample_CpG_stacked$method == methods[i]),c("depth")])),series = "observed")
  actual_CpG_cov$perc_CpG_SE_low <- NA
  actual_CpG_cov$perc_CpG_SE_high <- NA
  pred_act_cov <- rbind(predicted_CpG_cov, actual_CpG_cov) 
  pred_act_cov$method <- methods[i]
  predicted_CpG_cov_all_methods <- rbind(predicted_CpG_cov_all_methods,pred_act_cov)
}

#threshold Vmax to 1
predicted_CpG_cov_all_methods$perc_CpG_SE_high[predicted_CpG_cov_all_methods$perc_CpG_SE_high > 1] <- 1

model_df$coefficient <- rownames(model_df)
model_df$coefficient <- gsub("d.*","Vmax", model_df$coefficient)
model_df$coefficient <- gsub("e.*", "Km", model_df$coefficient)

pact_model_df <- model_df %>% tidyr::pivot_longer(names_to = "statistic", values_to = "value",1:6)%>% tidyr::pivot_wider(names_from = coefficient, values_from = value)

pact_model_df$species <- "P.acuta"
```

plot saturation curves per method
```{r}
#b <- ggplot(predicted_CpG_cov_all_methods) + geom_point(aes(x = depth, y = perc_CpG, color = method, alpha = series)) + scale_x_continuous(breaks = seq(0,4000,100)) + scale_y_continuous(labels = as.character(seq(0,1.0,0.2)),breaks = seq(0,1,0.2), limits = c(0,1),)+ theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.background =element_rect(fill="white"), plot.title =  element_text(face = "italic")) + ylab("fraction of genome-wide\nCpGs covered") + xlab("depth (M reads)") + labs(color = "method" ) + ggtitle("P. acuta") + scale_color_manual(values =c("#E6550D","#756BB1","#31A354")) + scale_alpha_manual(values=c(0.3, 1))


b <- ggplot(predicted_CpG_cov_all_methods,aes(x = depth, y = perc_CpG, color = method, alpha = series)) + geom_point(size = 0.5) + geom_line() + geom_ribbon(aes(ymin=perc_CpG_SE_low,ymax=perc_CpG_SE_high,fill=method),color="grey70",alpha=0.2, size = 0.1) + scale_x_continuous(breaks = seq(0,4000,100))+ scale_y_continuous(labels = as.character(seq(0,1.0,0.2)),breaks = seq(0,1,0.2)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.background =element_rect(fill="white"), plot.title =  element_text(face = "italic")) + ylab("fraction of genome-wide\nCpGs covered") + xlab("depth (M reads)") + labs(color = "method" ) + ggtitle("P. acuta") + scale_color_manual(values =c("#E6550D","#756BB1","#31A354")) + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354")) +  scale_alpha_manual(values=c(0.3, 1))

```

generate 2 panel figure
```{r}
jpeg("OLD_Output/Mcap_Pact_gnm-wide_CpGcov_estSatur.jpg", width = 8, height = 8, units = "in", res = 300)
ggarrange(a,b, labels = c("A", "B"),nrow = 2,common.legend = TRUE, legend = "top")
dev.off()
```


output model statistics for pact and mcap
```{r}
model_df_all <- rbind(mcap_model_df,pact_model_df)
write.csv(model_df_all, "Output/Seq_satur_est_CpGcov_Mcap_Pact.csv", quote = F, row.names = F)
```





TRY WITH FIXED VMAX; model coverage needed to saturate CpGs covered
```{r}
#create object that is a vector of the methods
methods <- unique(Mcap_pool_Sample_CpG_stacked$method)
#create empty df
predicted_CpG_cov_all_methods <- data.frame()
#create an empty data frame to save the model info in
model_df <- data.frame()
#loop through data 
for (i in 1:length(methods)){
  # use the two parameter MM model (MM.2)
  m1 <- drm(percent_CpG ~ as.integer(as.character(depth)), data =Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),], fct = MM.2(fixed = c(1, NA), names = c("d", "e")))
  #create vmax object from model
  Vmax <- 1
  #create K object from model
  Km <- data.frame(tibble(coef(m1)))[1,]
  #create model df to save model for each iteration
  mm <- data.frame(cbind(summary(m1)[[3]],summary(m1)[[13]]))
  #create a Vmax and Km SE high and low
  #Vmax_SE_low <- Vmax - mm$Std..Error[1]
  #Vmax_SE_high <- Vmax + mm$Std..Error[1]
  #Km_SE_low <- Km - mm$Std..Error[2]
  #Km_SE_high <- Km + mm$Std..Error[2]
  #add method column
  mm$method <- rep(methods[i],1)
  #combine model info with model info master df
  model_df <- rbind(model_df, mm)
  #create empty df
  predicted_CpG_cov <- data.frame()
  #create a series of numbers 50-3000 by 50s that will be the estimated depth
  prediction <- seq(50,4000, by = 50)
  #loop through the depths 
  for (j in 1:length(prediction)){
    #calculate the estimated percent CpGs covered (using MM model coefficients)
    perc_CpG <- (Vmax*prediction[j]/(Km+prediction[j]))
    # calculate perc_CpG for SE high and low
   # perc_CpG_SE_high <- (Vmax_SE_high*prediction[j]/(Km_SE_high+prediction[j]))
    #perc_CpG_SE_low <- (Vmax_SE_low*prediction[j]/(Km_SE_low+prediction[j]))
    #save the depth as an object
    depth <- prediction[j]
    #save the percent CpGs covered and the depth as a data frame
    pred_perc_CpG <- data.frame(perc_CpG,depth)
    #combine data frame for each depth in a master data frame
    predicted_CpG_cov <- rbind(predicted_CpG_cov,pred_perc_CpG)
  }
  #create column for denoting series
  #predicted_CpG_cov <- tidyr::gather(predicted_CpG_cov,"subseries", "perc_CpG", 1:3)
  #predicted_CpG_cov$subseries <- gsub("perc_CpG", "estimated", predicted_CpG_cov$subseries)
  predicted_CpG_cov$series <- "estimated"
  #create an object for the actual CpGs covered at known depths
  actual_CpG_cov <- data.frame(perc_CpG = Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),c("percent_CpG")],depth = as.integer(as.character(Mcap_pool_Sample_CpG_stacked[which(Mcap_pool_Sample_CpG_stacked$coverage==5 & Mcap_pool_Sample_CpG_stacked$method == methods[i]),c("depth")])),series = "observed")
  #actual_CpG_cov$subseries <- "observed"
  #actual_CpG_cov$perc_CpG_SE_low <- NA
  #actual_CpG_cov$perc_CpG_SE_high <- NA
  pred_act_cov <- rbind(predicted_CpG_cov, actual_CpG_cov) 
  pred_act_cov$method <- methods[i]
  predicted_CpG_cov_all_methods <- rbind(predicted_CpG_cov_all_methods,pred_act_cov)
}

model_df$coefficient <- rownames(model_df)
model_df$coefficient <- gsub("d.*","Vmax", model_df$coefficient)
model_df$coefficient <- gsub("e.*", "Km", model_df$coefficient)

mcap_model_df <- model_df %>% tidyr::pivot_longer(names_to = "statistic", values_to = "value",1:6)%>% tidyr::pivot_wider(names_from = coefficient, values_from = value)

mcap_model_df$species <- "M.capitata"

```


