---
title: "PCA for qualimap and meth profiles"
author: "Shelly Trigg"
date: "7/9/2020"
output: html_document
---


load libraries
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

read in union bed methylation data
```{r}
#read in union bedgraph for Mcap
Mcap <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Mcap_union_5x.bedgraph")

#read in union bedgraph for Pact
Pact <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Pact_union_5x.bedgraph")
```

convert N/A to NA
```{r}
Mcap[Mcap == "N/A"] <- NA
Pact[Pact == "N/A"] <- NA
```

filter for loci covered in all samples by removing lines with NA
```{r}
Mcap_complete <- data.frame(Mcap[complete.cases(Mcap),])
Pact_complete <- data.frame(Pact[complete.cases(Pact),])

```

see the structure of the df
```{r}
str(Mcap_complete)
str(Pact_complete)

```

The % methylation columns are character vectors because they initially contained "N/A" so convert all % methylation columns from character to numeric
```{r}
Mcap_complete[,-c(1:3)] <- sapply(Mcap_complete[,-c(1:3)], as.numeric)
Pact_complete[,-c(1:3)] <- sapply(Pact_complete[,-c(1:3)], as.numeric)

```

check the structure again
```{r}
str(Mcap_complete)
str(Pact_complete)

```
The conversion worked.

create rownames that is the loci ID (chrom:start-stop)
```{r}
row.names(Mcap_complete) <- paste0(Mcap_complete$chrom,":", Mcap_complete$start,"-",Mcap_complete$end)

row.names(Pact_complete) <- paste0(Pact_complete$chrom,":", Pact_complete$start,"-",Pact_complete$end)
```

remove chrom, start, and end columns
```{r}
Mcap_complete <- Mcap_complete[,-c(1:3)]
Pact_complete <- Pact_complete[,-c(1:3)]
```


swap sample numbers for sample names in colnames
```{r}
colnames(Mcap_complete) <- c("WGBS_1", "WGBS_2", "WGBS_3", "RRBS_1", "RRBS_2", "RRBS_3", "MBDBS_1", "MBDBS_2", "MBDBS_3")

colnames(Pact_complete) <- c("WGBS_1", "WGBS_2", "WGBS_3", "RRBS_1", "RRBS_2", "RRBS_3", "MBDBS_1", "MBDBS_2", "MBDBS_3")
```

write out input files for QC
```{r}
write.table(Mcap_complete, "OLD_Output/Mcap_union5xCpG_PCAinput.tsv", sep ="\t", quote = F)

write.table(Pact_complete, "OLD_Output/Pact_union5xCpG_PCAinput.tsv", sep ="\t", quote = F)

```

```{r}
Mcap_complete <- read.table("OLD_Output/Mcap_union5xCpG_PCAinput.tsv", sep = "\t", stringsAsFactors = F)

Pact_complete <-read.table("OLD_Output/Pact_union5xCpG_PCAinput.tsv", sep = "\t", stringsAsFactors = F)
```


create a meta data df
```{r}
Mcap_meta <- data.frame(row.names = colnames(Mcap_complete), Method = gsub("_.*","",colnames(Mcap_complete)))

Pact_meta <- data.frame(row.names = colnames(Pact_complete), Method = gsub("_.*","",colnames(Pact_complete)))
```

run PCA on union bed methylation data
```{r}
set.seed(101)
Mcap_pca <- prcomp(t(Mcap_complete))
Pact_pca <- prcomp(t(Pact_complete))
```

add method for plotting
```{r}
Mcap_pca_meta <- merge(Mcap_meta,Mcap_pca$x, by = "row.names")
Pact_pca_meta <- merge(Pact_meta,Pact_pca$x, by = "row.names")
```


plot PC 1 and 2 scores

```{r}
#PCA of union bed for Mcap
a <- ggplot(Mcap_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Mcap_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method") + ggtitle(expression(italic("M.capitata")))

#PCA of union bed for Pact
b <- ggplot(Pact_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Pact_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method") + ggtitle(expression(italic("P.acuta")))

figure <- ggarrange(a, b, labels = c("A", "B"), ncol = 2,common.legend = TRUE, legend = "bottom")

jpeg("OLD_Output/PCA_5xUnionBedCpG_Methylation.jpg", width = 8, height = 4, units = "in", res = 300)
figure
dev.off()

#adds sample names to plot
#ggplot(pact_pca_meta, aes(PC1, PC2, label = Sample)) + geom_point(aes(color = Group)) + geom_text(size = 1.5,hjust=1,vjust=-1) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + theme_bw() + ylab("PC2") + xlab("PC1")

```


