---
title: "PCA for qualimap and meth profiles"
author: "Shelly Trigg"
date: "7/9/2020"
output: html_document
---


load libraries
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

read in union bed methylation data
```{r}
#read in union bedgraph for Mcap
#Mcap <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Mcap_union_5x.bedgraph")

#read in union bedgraph for Pact
#Pact <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Pact_union_5x.bedgraph")


Mcap_5x <- read.table("output/intermediate-files/methPCA/Mcap_percmeth_5x.txt", header = T)
Mcap_10x <- read.table("output/intermediate-files/methPCA/Mcap_percmeth_10x.txt", header = T)
Mcap_50x <- read.table("output/intermediate-files/methPCA/Mcap_percmeth_50x.txt", header = T)
Mcap_100x <- read.table("output/intermediate-files/methPCA/Mcap_percmeth_100x.txt", header = T)

Pact_5x <- read.table("output/intermediate-files/methPCA/Pact_percmeth_5x.txt", header = T)
Pact_10x <- read.table("output/intermediate-files/methPCA/Pact_percmeth_10x.txt", header = T)
Pact_50x <- read.table("output/intermediate-files/methPCA/Pact_percmeth_50x.txt", header = T)
Pact_100x <- read.table("output/intermediate-files/methPCA/Pact_percmeth_100x.txt", header =T)




#read in union bedgraph for Pact
#Pact <- fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Pact_union_5x.bedgraph")


```

convert N/A to NA
```{r}
#Mcap[Mcap == "N/A"] <- NA
#Pact[Pact == "N/A"] <- NA
```

filter for loci covered in all samples by removing lines with NA
```{r}
#Mcap_complete <- data.frame(Mcap[complete.cases(Mcap),])
#Pact_complete <- data.frame(Pact[complete.cases(Pact),])

```

see the structure of the df
```{r}
#str(Mcap_complete)
#str(Pact_complete)

```

The % methylation columns are character vectors because they initially contained "N/A" so convert all % methylation columns from character to numeric
```{r}
#Mcap_complete[,-c(1:3)] <- sapply(Mcap_complete[,-c(1:3)], as.numeric)
#Pact_complete[,-c(1:3)] <- sapply(Pact_complete[,-c(1:3)], as.numeric)

```

check the structure again
```{r}
#str(Mcap_complete)
#str(Pact_complete)

```
The conversion worked.

create rownames that is the loci ID (chrom:start-stop)
```{r}
#row.names(Mcap_complete) <- paste0(Mcap_complete$chrom,":", Mcap_complete$start,"-",Mcap_complete$end)

#row.names(Pact_complete) <- paste0(Pact_complete$chrom,":", Pact_complete$start,"-",Pact_complete$end)
```

remove chrom, start, and end columns
```{r}
#Mcap_complete <- Mcap_complete[,-c(1:3)]
#Pact_complete <- Pact_complete[,-c(1:3)]
```


swap sample numbers for sample names in colnames
```{r}
#colnames(Mcap_complete) <- c("WGBS_1", "WGBS_2", "WGBS_3", "RRBS_1", "RRBS_2", "RRBS_3", "MBDBS_1", "MBDBS_2", "MBDBS_3")

#colnames(Pact_complete) <- c("WGBS_1", "WGBS_2", "WGBS_3", "RRBS_1", "RRBS_2", "RRBS_3", "MBDBS_1", "MBDBS_2", "MBDBS_3")
```

write out input files for QC
```{r}
#write.table(Mcap_complete, "OLD_Output/Mcap_union5xCpG_PCAinput.tsv", sep ="\t", quote = F)

#write.table(Pact_complete, "OLD_Output/Pact_union5xCpG_PCAinput.tsv", sep ="\t", quote = F)

```

```{r}
#Mcap_complete <- read.table("OLD_Output/Mcap_union5xCpG_PCAinput.tsv", sep = "\t", stringsAsFactors = F, header = T, row.names = 1)

#Pact_complete <-read.table("OLD_Output/Pact_union5xCpG_PCAinput.tsv", sep = "\t", stringsAsFactors = F,header = T, row.names = 1)
```


create a meta data df
```{r}
#Mcap_meta <- data.frame(row.names = colnames(Mcap_complete), Method = gsub("_.*","",colnames(Mcap_complete)))

Mcap_meta <- data.frame(row.names = colnames(Mcap_5x), Method = gsub("_.*","",colnames(Mcap_5x)))
Pact_meta <- data.frame(row.names = colnames(Pact_5x), Method = gsub("_.*","",colnames(Pact_5x)))


#Pact_meta <- data.frame(row.names = colnames(Pact_complete), Method = gsub("_.*","",colnames(Pact_complete)))
```

run PCA on union bed methylation data
```{r}
set.seed(101)
Mcap_5x_pca <- prcomp(t(Mcap_5x))
Mcap_10x_pca <- prcomp(t(Mcap_10x))
Mcap_50x_pca <- prcomp(t(Mcap_50x))
Mcap_100x_pca <- prcomp(t(Mcap_100x))


Pact_5x_pca <- prcomp(t(Pact_5x))
Pact_10x_pca <- prcomp(t(Pact_10x))
Pact_50x_pca <- prcomp(t(Pact_50x))
Pact_100x_pca <- prcomp(t(Pact_100x))

#Pact_pca <- prcomp(t(Pact_complete))
```

add method for plotting
```{r}
Mcap_5x_pca_meta <- merge(Mcap_meta,Mcap_5x_pca$x, by = "row.names")
Mcap_10x_pca_meta <- merge(Mcap_meta,Mcap_10x_pca$x, by = "row.names")
Mcap_50x_pca_meta <- merge(Mcap_meta,Mcap_50x_pca$x, by = "row.names")
Mcap_100x_pca_meta <- merge(Mcap_meta,Mcap_100x_pca$x, by = "row.names")

Pact_5x_pca_meta <- merge(Pact_meta,Pact_5x_pca$x, by = "row.names")
Pact_10x_pca_meta <- merge(Pact_meta,Pact_10x_pca$x, by = "row.names")
Pact_50x_pca_meta <- merge(Pact_meta,Pact_50x_pca$x, by = "row.names")
Pact_100x_pca_meta <- merge(Pact_meta,Pact_100x_pca$x, by = "row.names")

#add individual ID
Mcap_5x_pca_meta$Row.names <- gsub(".*_", "", Mcap_5x_pca_meta$Row.names)
Mcap_10x_pca_meta$Row.names <- gsub(".*_", "", Mcap_10x_pca_meta$Row.names)
Mcap_50x_pca_meta$Row.names <- gsub(".*_", "", Mcap_50x_pca_meta$Row.names)
Mcap_100x_pca_meta$Row.names <- gsub(".*_", "", Mcap_100x_pca_meta$Row.names)


Pact_5x_pca_meta$Row.names <- gsub(".*_", "", Pact_5x_pca_meta$Row.names)
Pact_10x_pca_meta$Row.names <- gsub(".*_", "", Pact_10x_pca_meta$Row.names)
Pact_50x_pca_meta$Row.names <- gsub(".*_", "", Pact_50x_pca_meta$Row.names)
Pact_100x_pca_meta$Row.names <- gsub(".*_", "", Pact_100x_pca_meta$Row.names)

#Pact_pca_meta <- merge(Pact_meta,Pact_pca$x, by = "row.names")
```


plot PC 1 and 2 scores

```{r}
#PCA of union bed for Mcap
#a <- ggplot(Mcap_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Mcap_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method") + ggtitle(expression(italic("M.capitata")))

a <- ggplot(Mcap_5x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_5x_pca)[[6]][[5]]*100,2),"% variation explained)"))  +xlab(paste0("PC1 (",round(summary(Mcap_5x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + ggtitle("read threshold: 5x") +  theme(plot.title = element_text(hjust = 0.5)) + annotate("text", x=-600, y=-550, label= paste0("n = ",nrow(Mcap_5x)))

b <- ggplot(Mcap_10x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_10x_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Mcap_10x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + ggtitle("read threshold: 10x") +  theme(plot.title = element_text(hjust = 0.5)) + annotate("text", x=500, y=-300, label= paste0("n = ",nrow(Mcap_10x)))

c <- ggplot(Mcap_50x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_50x_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Mcap_50x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + ggtitle("read threshold: 50x") +  theme(plot.title = element_text(hjust = 0.5)) + annotate("text", x=300, y=-150, label= paste0("n = ",nrow(Mcap_50x)))

d <- ggplot(Mcap_100x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Mcap_100x_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Mcap_100x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + ggtitle("read threshold: 100x") +  theme(plot.title = element_text(hjust = 0.5)) + annotate("text", x=175, y=-38, label= paste0("n = ",nrow(Mcap_100x)))

e <- ggplot(Pact_5x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_5x_pca)[[6]][[5]]*100,2),"% variation explained)")) +xlab(paste0("PC1 (",round(summary(Pact_5x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + annotate("text", x=-2800, y=-3500, label= paste0("n = ",nrow(Pact_5x)))

f <- ggplot(Pact_10x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method,shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_10x_pca)[[6]][[5]]*100,2),"% variation explained)")) +  xlab(paste0("PC1 (",round(summary(Pact_10x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + annotate("text", x=-2000, y=-1500, label= paste0("n = ",nrow(Pact_10x)))

g <- ggplot(Pact_50x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method, shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_50x_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Pact_50x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + annotate("text", x=-1000, y=-500, label= paste0("n = ",nrow(Pact_50x)))

h <- ggplot(Pact_100x_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method, shape = as.factor(Row.names)), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_100x_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Pact_100x_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method", shape = "individual") + annotate("text", x=-600, y=-250, label= paste0("n = ",nrow(Pact_100x)))

#PCA of union bed for Pact
#b <- ggplot(Pact_pca_meta, aes(PC1, PC2)) + geom_point(aes(color = Method), size = 3) + scale_color_manual(values = c("#FD8D3C","#9E9AC8","#74C476")) + theme_bw() + ylab(paste0("PC2 (", round(summary(Pact_pca)[[6]][[5]]*100,2),"% variation explained)")) + xlab(paste0("PC1 (",round(summary(Pact_pca)[[6]][[2]]*100,2),"% variation explained)")) + labs(color = "method") + ggtitle(expression(italic("P.acuta")))

figure <- ggarrange(a, b,c,d,e,f,g,h, labels = "AUTO",font.label = list(size = 18) , ncol = 4,nrow = 2,common.legend = TRUE, legend = "top")

jpeg("output/intermediate-files/methPCA/5x-100xCpG_Methylation.jpg", width = 15, height = 8, units = "in", res = 300)
figure
dev.off()

#adds sample names to plot
#ggplot(pact_pca_meta, aes(PC1, PC2, label = Sample)) + geom_point(aes(color = Group)) + geom_text(size = 1.5,hjust=1,vjust=-1) + scale_color_manual(values = c("#E6550D", "#756BB1","#31A354" )) + theme_bw() + ylab("PC2") + xlab("PC1")

```

