---
title: "Untitled"
author: "Shelly Trigg"
date: "8/28/2020"
output: html_document
---


load libraries
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tidyr)
```

read in union bed methylation data
```{r}
#read in union bedgraph for Mcap
Mcap <- data.frame(fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Mcap_union_5x.bedgraph"))

#read in union bedgraph for Pact
Pact <- data.frame(fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200424/10-unionbedg/Pact_union_5x.bedgraph"))
```

reformat data
```{r}
#convert to long
Mcap_l <- tidyr::gather(data.frame(Mcap),"Sample","perc_meth",4:12)
Pact_l <- tidyr::gather(data.frame(Pact),"Sample","perc_meth",4:12)
```

convert N/A to NA
```{r}
Mcap_l[Mcap_l == "N/A"] <- NA
Pact_l[Pact_l == "N/A"] <- NA
```

remove rows with NA
```{r}
Mcap_l <- data.frame(Mcap_l[complete.cases(Mcap_l),])
Pact_l <- data.frame(Pact_l[complete.cases(Pact_l),])
```

create column that is the loci ID (chrom:start-stop)
```{r}
Mcap_l$loci <- paste0(Mcap_l$chrom,":", Mcap_l$start,"-",Mcap_l$end)

Pact_l$loci <- paste0(Pact_l$chrom,":", Pact_l$start,"-",Pact_l$end)
```

remove columns 1-3
```{r}
Mcap_l <- Mcap_l[,-c(1:3)]
Pact_l <- Pact_l[,-c(1:3)]
```


convert sample column and perc meth column to numeric
```{r}
Mcap_l$Sample <- gsub("X","", Mcap_l$Sample)
Pact_l$Sample <- gsub("X","", Pact_l$Sample)

Mcap_l$Sample <- as.numeric(as.character(Mcap_l$Sample))
Pact_l$Sample <- as.numeric(as.character(Pact_l$Sample))

#The % methylation column are character vectors because they initially contained "N/A" so convert all % methylation columns from character to numeric

Mcap_l$perc_meth <- as.numeric(as.character(Mcap_l$perc_meth))
Pact_l$perc_meth <- as.numeric(as.character(Pact_l$perc_meth))
```

Add method labels
```{r}
Mcap_l$method <- gsub("10|11|12","WGBS",Mcap_l$Sample)
Mcap_l$method <- gsub("13|14|15","RRBS",Mcap_l$Sample)
Mcap_l$method <- gsub("16|17|18","RRBS",Mcap_l$Sample)

Pact_l$method <- Pact_l$Sample
Pact_l$method <- gsub("1|2|3","WGBS",Pact_l$method)
Pact_l$method <- gsub("4|5|6","RRBS",Pact_l$method)
Pact_l$method <- gsub("7|8|9","MBD",Pact_l$method)

```

Create group column with methods listed as string
```{r}
Pact_loci_method <- Pact_l %>% dplyr::group_by(loci) %>% dplyr::summarise(group = toString(method))

table(Pact_loci_method$group)

#remove duplicate methods (e.g. convert RRBS, RRBS, RRBS to RRBS)
#create a short test dataset
tail.Pact_l <- tail(Pact_l, 100)
tail.Pact_l <- tail.Pact_l %>% dplyr::group_by(loci) %>% dplyr::summarise(group = toString(method))
#create function to convert methods for common loci into a string then split the methods string into a character vector, and use unique function to keep only unique characters
uniqMethod.FUN <- function(x){paste(sort(unique(unlist(strsplit(x, split =", "), use.names = F))), collapse = ", ")}
#apply the function to each row in the df, and save the output in a new column
tail.Pact_l$group_simp <- sapply(tail.Pact_l$group, uniqMethod.FUN, USE.NAMES = F)
str(tail.Pact_l)
# the function works so apply to full df
table(tail.Pact_l$group_simp)
#looks good so apply to full df
Pact_loci_method$group_simp <- sapply(Pact_loci_method$group, uniqMethod.FUN, USE.NAMES = F)

#check df
str(Pact_loci_method)
table(Pact_loci_method$group_simp)

#combine back with main table
nrow(Pact_loci_method)
nrow(Pact_l)

Pact_l_m <- merge(Pact_loci_method,Pact_l, by = "loci")
nrow(Pact_l_m)
str(Pact_l_m)
```


make bar plot of data
```{r}

##binning data
## found this useful: https://www.jdatalab.com/data_science_and_data_mining/2017/01/30/data-binning-plot.html

## add bin column
# specify interval/bin labels
tags <- c("[0-10]","[10-20]", "[20-30]", "[30-40]", "[40-50]", "[50-60]", "[60-70]", "[70-80]","[80-90]", "[90-100]")

# bucketing values into bins

Pact_l_m_tag <- as_tibble(Pact_l_m) %>% mutate(tag = case_when(perc_meth < 10 ~ tags[1],perc_meth >= 10 & perc_meth < 20 ~ tags[2],perc_meth >= 20 & perc_meth < 30 ~ tags[3],perc_meth >= 30 & perc_meth < 40 ~ tags[4],perc_meth >= 40 & perc_meth < 50 ~ tags[5], perc_meth >= 50 & perc_meth < 60 ~ tags[6], perc_meth >= 60 & perc_meth < 70 ~ tags[7], perc_meth >= 70 & perc_meth < 80 ~ tags[8], perc_meth >= 80 & perc_meth < 90 ~ tags[9], perc_meth >= 90 & perc_meth <= 100 ~ tags[10]))

#order group_simp
Pact_l_m_tag$group_simp <- factor(Pact_l_m_tag$group_simp, levels = c("MBD", "RRBS", "WGBS", "MBD, RRBS","MBD, WGBS", "RRBS, WGBS", "MBD, RRBS, WGBS"))

#simplify df for plotting; this creates a shorter df
#followed these: 

# plot percentage instead of counts
#https://stackoverflow.com/questions/24776200/ggplot-replace-count-with-percentage-in-geom-bar/48602277

#add num observations to top of plot
#https://stackoverflow.com/questions/46984296/proportion-with-ggplot-geom-bar

#expand space between facet label and plot area
#https://www.datanovia.com/en/blog/ggplot-facet-how-to-add-space-between-labels-on-the-top-of-the-chart-and-the-plot-border/

Pact_l_m_tag_STACKED <- Pact_l_m_tag %>% group_by(group_simp, method, tag) %>% summarise(count=n()) %>% 
  mutate(perc_CpG=count/sum(count), cumsum = cumsum(perc_CpG), label = ifelse(tag=="[90-100]",paste0("N=", sum(count)),""))


jpeg("OLD_Output/Pact_5xCpGmeth_overlapMethods_StackedBars_colbyMethod.jpg", width = 8, height = 8, units = "in", res = 300)
ggplot(Pact_l_m_tag_STACKED,aes(x = method, y = perc_CpG*100, fill = as.factor(method), alpha = tag, group = method )) + geom_bar(stat = "identity",position = "fill", color = "black") + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354"))  + scale_y_continuous(expand = expansion(mult = c(0, 0.1)),labels=scales::percent) + scale_alpha_manual(values=c(seq(0.1,1, length.out = 10)))+ labs(x = "method", y = "% CpGs",fill = "method", alpha = "% methylation") + facet_wrap(~group_simp) +  geom_text(aes(y=cumsum, label=label), vjust=-0.5, size = 2.5) + guides(fill=guide_legend(reverse=T)) + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
dev.off()



jpeg("OLD_Output/Pact_5xCpGmeth_overlapMethods_StackedBars.jpg", width = 8, height = 8, units = "in", res = 300)
ggplot(Pact_l_m_tag,aes(x = group_simp, y = ..count.., fill = group_simp, alpha = tag )) + geom_bar(position = "fill", color = "black") + scale_y_continuous(labels=scales::percent) + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + scale_fill_manual(values =c("red","red","red", "blue", "blue","blue", "black"))  + scale_alpha_manual(values=c(seq(0.1,1, length.out = 10)))+ labs(x = "method", y = "% CpGs",fill = "method", alpha = "% methylation")
dev.off()


#histograms
jpeg("OLD_Output/Pact_5xCpGmeth_overlapMethods_Histograms.jpg", width = 8, height = 8, units = "in", res = 300)
ggplot(Pact_l_m_tag,aes(x = perc_meth, fill = group_simp,group = method, color = "black")) + geom_histogram(position = "dodge",binwidth = 10) + ylab("num CpGs") + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + facet_wrap(~group_simp, scale = "free") + xlab("% methylation") + scale_color_manual(values = "black") + scale_fill_manual(values =c("red","red","red", "blue", "blue","blue", "black")) + theme(legend.position="none")
dev.off()

jpeg("OLD_Output/Pact_5xCpGmeth_overlapMethods_Histograms_colbyMethod.jpg", width = 8, height = 8, units = "in", res = 300)
ggplot(Pact_l_m_tag,aes(x = perc_meth, fill = method,group = method, color = "black")) + geom_histogram(position = "dodge",binwidth = 10) + ylab("num CpGs") + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + facet_wrap(~group_simp, scale = "free") + xlab("% methylation") + scale_color_manual(values = "black") + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354"))
dev.off()


```

This chuck is just a test I did to figure out the plotting parameters 
```{r}

tail.Pact_l_m <- tail(Pact_l_m,100000)

#add column for simplifying overlap groups
tail.Pact_l_m$group_more_simp <- tail.Pact_l_m$group_simp
tail.Pact_l_m$group_more_simp <- gsub("^MBD$|^RRBS$|^WGBS$", "1.method",tail.Pact_l_m$group_more_simp)
tail.Pact_l_m$group_more_simp <- gsub("^MBD, WGBS$|^RRBS, WGBS$", "2.methods",tail.Pact_l_m$group_more_simp)
tail.Pact_l_m$group_more_simp <- gsub("^MBD, RRBS, WGBS$", "3.methods",tail.Pact_l_m$group_more_simp)

## add bin column

# set up cut-off values 
breaks <- seq(0,100,by = 10)
# specify interval/bin labels
tags <- c("[0-10]","[10-20]", "[20-30]", "[30-40]", "[40-50]", "[50-60]", "[60-70]", "[70-80]","[80-90]", "[90-100]")

# bucketing values into bins

tail.Pact_l_m_tag <- as_tibble(tail.Pact_l_m) %>% mutate(tag = case_when(perc_meth < 10 ~ tags[1],perc_meth >= 10 & perc_meth < 20 ~ tags[2],perc_meth >= 20 & perc_meth < 30 ~ tags[3],perc_meth >= 30 & perc_meth < 40 ~ tags[4],perc_meth >= 40 & perc_meth < 50 ~ tags[5], perc_meth >= 50 & perc_meth < 60 ~ tags[6], perc_meth >= 60 & perc_meth < 70 ~ tags[7], perc_meth >= 70 & perc_meth < 80 ~ tags[8], perc_meth >= 80 & perc_meth < 90 ~ tags[9], perc_meth >= 90 & perc_meth <= 100 ~ tags[10]))

#order group_simp
tail.Pact_l_m_tag$group_simp <- factor(tail.Pact_l_m$group_simp, levels = c("MBD", "RRBS", "WGBS", "MBD, RRBS","MBD, WGBS", "RRBS, WGBS", "MBD, RRBS, WGBS"))

tail.Pact_l_m_tag_STACKED <- tail.Pact_l_m_tag %>% group_by(group_simp, method, tag) %>% summarise(count=n()) %>% 
  mutate(perc_CpG=count/sum(count), cumsum = cumsum(perc_CpG), label = ifelse(tag=="[90-100]",paste0("N=", sum(count)),""))


ggplot(tail.Pact_l_m_tag_STACKED,aes(x = method, y = perc_CpG*100, fill = as.factor(method), alpha = tag, group = method )) + geom_bar(stat = "identity",position = "fill", color = "black") + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354"))  + scale_y_continuous(expand = expansion(mult = c(0, 0.1)),labels=scales::percent) + scale_alpha_manual(values=c(seq(0.1,1, length.out = 10)))+ labs(x = "method", y = "% CpGs",fill = "method", alpha = "% methylation") + facet_wrap(~group_simp) +  geom_text(aes(y=cumsum, label=label), vjust=-0.5, size = 3) + guides(fill=guide_legend(reverse=T))



#histograms
ggplot(tail.Pact_l_m_tag,aes(x = perc_meth, fill = method, group = method, color = "black")) + geom_histogram(position = "dodge",binwidth = 10) + ylab("num CpGs") + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + facet_wrap(~group_simp, scale = "free") + xlab("% methylation") + scale_color_manual(values = c("black")) + scale_fill_manual(values =c("#E6550D","#756BB1","#31A354")) + theme(legend.position="none")
```

