---
title: "UpsetPlot"
author: "Shelly Trigg"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

install packages
```{r}
install.packages("UpSetR")

if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("ComplexHeatmap")

```

load libraries
```{r}
library(UpSetR)
library(ComplexHeatmap)
library(data.table)
```

load data
```{r}
pact <- read.table("/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Pact/Pact_method_union_5x_matrix.tsv", header = TRUE, sep = "\t", row.names = 1, stringsAsFactors = FALSE)
  #fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200501/data/Pact_method_union_5x_matrix.tsv", stringsAsFactors = FALSE, header = TRUE)

mcap <- read.table("/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Mcap/Mcap_method_union_5x_matrix.tsv", header = TRUE, sep = "\t", row.names = 1, stringsAsFactors = FALSE)
```

change column name from "all" to "Genome"
```{r}
colnames(pact) <- c("Genome", "WGBS", "RRBS", "MBD")

colnames(mcap) <- c("Genome", "WGBS", "RRBS", "MBD")


```

make combination matrix (this is only necessary to produce the draft 1 plots)
```{r}
pact_m <- make_comb_mat(pact)
mcap_m <- make_comb_mat(mcap)

save(pact_m, file = "/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Pact/pact_m.RData")

save(mcap_m, file = "/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Mcap/mcap_m.RData")
```

Draft 1
```{r}
jpeg("../../Output/Pact_UpsetPlot.jpg", width = 8, height = 4, unit = "in", res = 300)
UpSet(pact_m,pt_size = unit(5, "mm"), lwd = 3,set_order = c("Genome", "WGBS", "RRBS", "MBD"), row_names_side = "right",comb_col = c("gold","red", "blue", "black")[comb_degree(pact_m)])
dev.off()

jpeg("../../Output/Mcap_UpsetPlot.jpg", width = 8, height = 4, unit = "in", res = 300)
UpSet(mcap_m,pt_size = unit(5, "mm"), lwd = 3,set_order = c("Genome", "WGBS", "RRBS", "MBD"),comb_col = c("gold","red", "blue", "black")[comb_degree(mcap_m)])
dev.off()
```

**Draft 2**
I changed the dot matrix colors to match the set color, made the main bars black, added numbers to all bars, ordered the main bars by sets, changed the axis labels, and offset the set bar size numbers.

To change the color of the dots, I edited the source code by doing trace("upset", edit=TRUE)
and followed edits from this post: https://www.r-bloggers.com/hacking-our-way-through-upsetr/

I basically just added the code below just after the Matrix_layout variable is assigned.
```
for(i in 1:length(unique(Matrix_layout$y))) {
     j <- which(Matrix_layout$y == i & Matrix_layout$value == 1)
     if(length(j) > 0) Matrix_layout$color[j] <- c("white", "springgreen3","dodgerblue", "red")[i]
  } 
```
An alternative for future plots so that the number of colors matches the number of set groups could be the following:
```
  for(i in 1:length(unique(Matrix_layout$y))) {
     j <- which(Matrix_layout$y == i & Matrix_layout$value == 1)
     if(length(j) > 0) Matrix_layout$color[j] <- colorRamps::matlab.like2(n = length(unique(Matrix_layout$y)))[i]
  }
```
To offset the set size bar numbers, I edited the source code trace(UpSetR:::Make_size_plot, edit = TRUE) hjust parameter to from 0.1 to 1 so that the set size number does not get cut off
```
if (set_size.show == TRUE) {
  Size_plot <- (Size_plot + geom_text(aes(label = y, vjust = 0.5, hjust = 0.3), angle = 45, size = (7/ggplot2:::.pt) * x_axis_tick_label_scale))
}
```
```{r}
jpeg("../../Output/Pact_UpsetPlot_draft2.jpg", width = 11, height = 7, unit = "in", res = 300)
upset(pact, sets = c("Genome", "WGBS", "MBD","RRBS"), keep.order = TRUE,group.by = "sets",main.bar.color = "black", sets.bar.color = c("orange", "springgreen3","dodgerblue", "red"), shade.alpha = 0.01,point.size = 4, line.size = 0.001, show.numbers = "yes",matrix.dot.alpha = 0.01, set_size.angles = 20, set_size.show = TRUE, mainbar.y.label = "Number of unique CpGs between sets", sets.x.label = "Set Size (number of CpGs)")
dev.off()

jpeg("../../Output/Mcap_UpsetPlot_draft2.jpg", width = 11, height = 7, unit = "in", res = 300)
upset(mcap, sets = c("Genome", "WGBS", "MBD","RRBS"), keep.order = TRUE, group.by = "sets",main.bar.color = "black", sets.bar.color = c("orange", "springgreen3","dodgerblue", "red"), shade.alpha = 0.01,point.size = 4, line.size = 0.001, show.numbers = "yes",matrix.dot.alpha = 0.01, set_size.angles = 20, set_size.show = TRUE, mainbar.y.label = "Number of unique CpGs between sets", sets.x.label = "Set Size (number of CpGs)")
dev.off()

```