---
title: "UpsetPlot"
author: "Shelly Trigg"
date: "5/1/2020"
output: html_document
---

install packages
```{r}
install.packages("UpSetR")

if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("ComplexHeatmap")

```

load libraries
```{r}
library(UpSetR)
library(ComplexHeatmap)
library(data.table)
```

load data
```{r}
pact <- read.table("/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Pact/Pact_method_union_5x_matrix.tsv", header = TRUE, sep = "\t", row.names = 1, stringsAsFactors = FALSE)
  #fread("https://gannet.fish.washington.edu/metacarcinus/FROGER_meth_compare/20200501/data/Pact_method_union_5x_matrix.tsv", stringsAsFactors = FALSE, header = TRUE)

mcap <- read.table("/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Mcap/Mcap_method_union_5x_matrix.tsv", header = TRUE, sep = "\t", row.names = 1, stringsAsFactors = FALSE)
```

change column name from "all" to "Genome"
```{r}
colnames(pact) <- c("Genome", "WGBS", "RRBS", "MBD")

colnames(mcap) <- c("Genome", "WGBS", "RRBS", "MBD")


```

make combination matrix
```{r}
pact_m <- make_comb_mat(pact)
mcap_m <- make_comb_mat(mcap)

save(pact_m, file = "/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Pact/pact_m.RData")

save(mcap_m, file = "/Volumes/web/metacarcinus/FROGER_meth_compare/20200504/data/Mcap/mcap_m.RData")
```

Draft 1
```{r}
jpeg("../../Output/Pact_UpsetPlot.jpg", width = 8, height = 4, unit = "in", res = 300)
UpSet(pact_m,pt_size = unit(5, "mm"), lwd = 3,set_order = c("Genome", "WGBS", "RRBS", "MBD"), row_names_side = "right",comb_col = c("gold","red", "blue", "black")[comb_degree(pact_m)])
dev.off()

jpeg("../../Output/Mcap_UpsetPlot.jpg", width = 8, height = 4, unit = "in", res = 300)
UpSet(mcap_m,pt_size = unit(5, "mm"), lwd = 3,set_order = c("Genome", "WGBS", "RRBS", "MBD"),comb_col = c("gold","red", "blue", "black")[comb_degree(mcap_m)])
dev.off()
```

**Draft 2**
To change the color of the dots, I editted the source code by doing trace("upset", edit=TRUE)
and followed edits from this post: https://www.r-bloggers.com/hacking-our-way-through-upsetr/

I basically just added the code below just after the Matrix_layout variable is assigned.
```
for(i in 1:length(unique(Matrix_layout$y))) {
     j <- which(Matrix_layout$y == i & Matrix_layout$value == 1)
     if(length(j) > 0) Matrix_layout$color[j] <- c("orange", "springgreen3","dodgerblue", "red")[i]
  } 
```
An alternative for future plots so that the number of colors matches the number of set groups could be the following:
```
  for(i in 1:length(unique(Matrix_layout$y))) {
     j <- which(Matrix_layout$y == i & Matrix_layout$value == 1)
     if(length(j) > 0) Matrix_layout$color[j] <- colorRamps::matlab.like2(n = length(unique(Matrix_layout$y)))[i]
  }
```

```{r}
jpeg("../../Output/Pact_UpsetPlot_draft2.jpg", width = 8, height = 4, unit = "in", res = 300)
upset(pact, sets = c("Genome", "WGBS", "MBD","RRBS"), keep.order = TRUE, order.by = "freq",main.bar.color = "black", sets.bar.color = c("orange", "springgreen3","dodgerblue", "red"), shade.alpha = 0.01,point.size = 4, line.size = 0.001, show.numbers = "yes",matrix.dot.alpha = 0.01, set_size.angles = 20, set_size.show = TRUE, mainbar.y.label = "Intersection Size (number of CpGs)", sets.x.label = "Set Size (number of CpGs)")
dev.off()

jpeg("../../Output/Mcap_UpsetPlot_draft2.jpg", width = 8, height = 4, unit = "in", res = 300)
upset(mcap, sets = c("Genome", "WGBS", "MBD","RRBS"), keep.order = TRUE, order.by = "freq",main.bar.color = "black", sets.bar.color = c("orange", "springgreen3","dodgerblue", "red"), shade.alpha = 0.01,point.size = 4, line.size = 0.001, show.numbers = "yes",matrix.dot.alpha = 0.01, set_size.angles = 20, set_size.show = TRUE, mainbar.y.label = "Intersection Size (number of CpGs)", sets.x.label = "Set Size (number of CpGs)")
dev.off()

```