---
title: "Chacterizing CpG Methylation (5x Individual Samples)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script, I'll create summary tables and figures to characterize CpG methylation in *M. capitata* and *P. acuta* using WGBS, RRBS, and MBD-BSSeq. I used individual samples with data for 5x CpGs in [this Jupyter notebook](https://github.com/hputnam/Meth_Compare/blob/master/scripts/Characterizing-CpG-Methylation-5x.ipynb) to identify methylation status and genomic location. I will use the output in this script.

# Install packages

```{r}
#install.packages("RColorBrewer")
require(RColorBrewer) #Load RColorBrewer

#install.packages("dichromat")
require(dichromat)
```

# Session information

```{r}
sessionInfo()
```

# CpG distribution

## Mcap

### Import file counts

```{r}
McapAll <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-bedgraph-counts.txt", header = FALSE, col.names = c("totalLines", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapAll <- McapAll[-10,] #Remove last row (total lines for all files)
tail(McapAll) #Confirm import
```

```{r}
McapMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-Meth-counts.txt", header = FALSE, col.names = c("Meth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapMeth <- McapMeth[-10,] #Remove last row (total lines for all files)
tail(McapMeth) #Confirm import
```

```{r}
McapSparseMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-sparseMeth-counts.txt", header = FALSE, col.names = c("sparseMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapSparseMeth <- McapSparseMeth[-10,] #Remove last row (total lines for all files)
tail(McapSparseMeth) #Confirm import
```

```{r}
McapUnMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-unMeth-counts.txt", header = FALSE, col.names = c("unMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapUnMeth <- McapUnMeth[-10,] #Remove last row (total lines for all files)
tail(McapUnMeth) #Confirm import
```

### Create summary table

```{r}
McapCpGType <- cbind(McapAll, McapMeth, McapSparseMeth, McapUnMeth) #Mash tables together by column
rownames(McapCpGType) <- substr(McapAll$filename, start = 1, stop = 6) #Use the first 6 characters of the filename to add sample ID as row names
McapCpGType <- McapCpGType[,-c(2,4,6,8)] #Remove filename columns
tail(McapCpGType) #Confirm table mashing
```

```{r}
McapCpGType$percentMeth <- (McapCpGType$Meth / McapCpGType$totalLines) * 100 #Calculate percent methylated loci
McapCpGType$percentSparseMeth <- (McapCpGType$sparseMeth / McapCpGType$totalLines) * 100 #Calculate percent sparsely methylated loci
McapCpGType$percentUnMeth <- (McapCpGType$unMeth / McapCpGType$totalLines) * 100 #Calculate percent unmethylated loci
McapCpGType <- McapCpGType[,c(1, 2, 5, 3, 6, 4, 7)] #Reorganize columns
tail(McapCpGType) #Confirm calculations
```

```{r}
write.table(McapCpGType, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CpG-Type.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

### Reorganize data

```{r}
McapCpGTypePercents <- McapCpGType[,c(3,5,7)] #Keep only columns with % total CpG information
head(McapCpGTypePercents) #Check reorganization
```

```{r}
#Create test plots

barplot(t(McapCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(McapCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, RRBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(McapCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, MBD-BSSeq
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")
```

### Create multipanel figure with all samples

```{r}
#pdf("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CpG-Type.pdf", height = 8.5, width = 11) #Save file as pdf

par(mfcol = c(3,3), mar = c(2, 2, 2, 0), oma = c(5, 5, 2, 0)) #Fill in multipanel plot by column and adjust inner and outer margins

barplot(t(McapCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 3, "WBGS", adj = 0, line = 1) #Add sequencing information

barplot(t(McapCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, RRBS
mtext(side = 3, "RRBS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 2, outer = TRUE, "% 5x CpG with Data", cex = 1.5, line = 2) #Add y-axis label

barplot(t(McapCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 1, MBD-BSSeq
mtext(side = 3, "MBD-BSSeq", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(McapCpGTypePercents[2,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, WGBS
barplot(t(McapCpGTypePercents[5,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, RRBS
barplot(t(McapCpGTypePercents[8,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 2, MBD-BSSeq
mtext(side = 1, outer = TRUE, "% Methylation", cex = 1.5, line = 2) #Add x-axis label

barplot(t(McapCpGTypePercents[3,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, WGBS
barplot(t(McapCpGTypePercents[6,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, RRBS
barplot(t(McapCpGTypePercents[9,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 3, MBD-BSSeq

#dev.off() #Turn off plotting device
```

## Pact

### Import file counts

```{r}
PactAll <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-bedgraph-counts.txt", header = FALSE, col.names = c("totalLines", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactAll <- PactAll[-10,] #Remove last row (total lines for all files)
tail(PactAll) #Confirm import
```

```{r}
PactMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-Meth-counts.txt", header = FALSE, col.names = c("Meth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactMeth <- PactMeth[-10,] #Remove last row (total lines for all files)
tail(PactMeth) #Confirm import
```

```{r}
PactSparseMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-sparseMeth-counts.txt", header = FALSE, col.names = c("sparseMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactSparseMeth <- PactSparseMeth[-10,] #Remove last row (total lines for all files)
tail(PactSparseMeth) #Confirm import
```

```{r}
PactUnMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-unMeth-counts.txt", header = FALSE, col.names = c("unMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactUnMeth <- PactUnMeth[-10,] #Remove last row (total lines for all files)
tail(PactUnMeth) #Confirm import
```

### Create summary table

```{r}
PactCpGType <- cbind(PactAll, PactMeth, PactSparseMeth, PactUnMeth) #Mash tables together by column
rownames(PactCpGType) <- substr(PactAll$filename, start = 1, stop = 5) #Use the first 5 characters of the filename to add sample ID to row names
PactCpGType <- PactCpGType[,-c(2,4,6,8)] #Remove filename columns
tail(PactCpGType) #Confirm table mashing
```

```{r}
PactCpGType$percentMeth <- (PactCpGType$Meth / PactCpGType$totalLines) * 100 #Calculate percent methylated loci
PactCpGType$percentSparseMeth <- (PactCpGType$sparseMeth / PactCpGType$totalLines) * 100 #Calculate percent sparsely methylated loci
PactCpGType$percentUnMeth <- (PactCpGType$unMeth / PactCpGType$totalLines) * 100 #Calculate percent unmethylated loci
PactCpGType <- PactCpGType[,c(1, 2, 5, 3, 6, 4, 7)] #Reorganize columns
tail(PactCpGType) #Confirm calculations
```

```{r}
write.table(PactCpGType, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CpG-Type.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

```{r}
PactCpGType <- read.delim("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CpG-Type.txt", sep = "\t", header = TRUE) #Import summary table
head(PactCpGType) #Check import
```

### Reorganize data

```{r}
PactCpGTypePercents <- PactCpGType[,c(3,5,7)] #Keep only columns with % total CpG information
tail(PactCpGTypePercents) #Check reorganization
```

```{r}
#Create test plots

barplot(t(PactCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(PactCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, RRBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(PactCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)")), axes = FALSE) #Sample 1, MBD-BSSeqc(expression("High (">="50%)"), "Moderate (10-50%)", expression("Weak ("<="10%)"))
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")
```

### Create multipanel figure with all samples

```{r}
#pdf("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CpG-Type.pdf", height = 8.5, width = 11) #Save file as pdf

par(mfcol = c(3,3), mar = c(2, 2, 2, 0), oma = c(5, 5, 2, 0)) #Fill in multipanel plot by column and adjust inner and outer margins

barplot(t(PactCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, WGBS
mtext(side = 3, "WBGS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(PactCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, RRBS
mtext(side = 3, "RRBS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 2, outer = TRUE, "% 5x CpG with Data", cex = 1.5, line = 2) #Add y-axis label

barplot(t(PactCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 1, MBD-BSSeq
mtext(side = 3, "MBD-BSSeq", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(PactCpGTypePercents[2,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, WGBS
barplot(t(PactCpGTypePercents[5,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, RRBS
barplot(t(PactCpGTypePercents[8,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 2, MBD-BSSeq
mtext(side = 1, outer = TRUE, "% Methylation", cex = 1.5, line = 2) #Add x-axis label

barplot(t(PactCpGTypePercents[3,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, WGBS
barplot(t(PactCpGTypePercents[6,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, RRBS
barplot(t(PactCpGTypePercents[9,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 3, MBD-BSSeq

#dev.off() #Turn off plotting device
```

# CpG overlaps

## Mcap

### Import file counts

```{r}
McapGenomeFeatures <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CGMotif-Overlaps-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with CG motif and feature track overlaps
McapGenomeFeatures <- McapGenomeFeatures[-8,] #Remove final row
tail(McapGenomeFeatures) #Check import
```

```{r}
McapGeneOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcGenes-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-gene overlaps
McapGeneOverlaps <- McapGeneOverlaps[-37,] #Remove final row
tail(McapGeneOverlaps) #Confirm import
```

```{r}
McapCDSOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcCDS-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-CDS overlaps
McapCDSOverlaps <- McapCDSOverlaps[-37,] #Remove final row
tail(McapCDSOverlaps) #Confirm import
```

```{r}
McapIntronsOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcIntrons-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Introns overlaps
McapIntronsOverlaps <- McapIntronsOverlaps[-37,] #Remove final row
tail(McapIntronsOverlaps) #Confirm import
```

```{r}
McapFlanksOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcFlanks-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-flank overlaps
McapFlanksOverlaps <- McapFlanksOverlaps[-37,] #Remove final row
tail(McapFlanksOverlaps) #Confirm import
```

```{r}
McapFlanksUpstreamOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcFlanksUpstream-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-upstream flank overlaps
McapFlanksUpstreamOverlaps <- McapFlanksUpstreamOverlaps[-37,] #Remove final row
tail(McapFlanksUpstreamOverlaps) #Confirm import
```

```{r}
McapFlanksDownstreamOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcFlanksDownstream-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Introns overlaps
McapFlanksUpstreamOverlaps <- McapFlanksUpstreamOverlaps[-37,] #Remove final row
tail(McapFlanksDownstreamOverlaps) #Confirm import
```

```{r}
McapIntergenicOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcIntergenic-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Intergenic overlaps
McapIntergenicOverlaps <- McapIntergenicOverlaps[-37,] #Remove final row
tail(McapIntergenicOverlaps) #Confirm import
```

### Create summary table

```{r}
McapFeatureOverlaps <- data.frame("allCpGs" = rep(0, times = 7),
                                  "Meth10Meth" = rep(0, times = 7),
                                  "Meth10sparseMeth" = rep(0, times = 7),
                                  "Meth10unMeth" = rep(0, times = 7),
                                  "Meth10" = rep(0, times = 7),
                                  "Meth11Meth" = rep(0, times = 7),
                                  "Meth11sparseMeth" = rep(0, times = 7),
                                  "Meth11unMeth" = rep(0, times = 7),
                                  "Meth11" = rep(0, times = 7),
                                  "Meth12Meth" = rep(0, times = 7),
                                  "Meth12sparseMeth" = rep(0, times = 7),
                                  "Meth12unMeth" = rep(0, times = 7),
                                  "Meth12" = rep(0, times = 7),
                                  "Meth13Meth" = rep(0, times = 7),
                                  "Meth13sparseMeth" = rep(0, times = 7),
                                  "Meth13unMeth" = rep(0, times = 7),
                                  "Meth13" = rep(0, times = 7),
                                  "Meth14Meth" = rep(0, times = 7),
                                  "Meth14sparseMeth" = rep(0, times = 7),
                                  "Meth14unMeth" = rep(0, times = 7),
                                  "Meth14" = rep(0, times = 7),
                                  "Meth15Meth" = rep(0, times = 7),
                                  "Meth15sparseMeth" = rep(0, times = 7),
                                  "Meth15unMeth" = rep(0, times = 7),
                                  "Meth15" = rep(0, times = 7),
                                  "Meth16Meth" = rep(0, times = 7),
                                  "Meth16sparseMeth" = rep(0, times = 7),
                                  "Meth16unMeth" = rep(0, times = 7),
                                  "Meth16" = rep(0, times = 7),
                                  "Meth17Meth" = rep(0, times = 7),
                                  "Meth17sparseMeth" = rep(0, times = 7),
                                  "Meth17unMeth" = rep(0, times = 7),
                                  "Meth17" = rep(0, times = 7),
                                  "Meth18Meth" = rep(0, times = 7),
                                  "Meth18sparseMeth" = rep(0, times = 7),
                                  "Meth18unMeth" = rep(0, times = 7),
                                  "Meth18" = rep(0, times = 7)) #Create blank dataframe with information for various CpG categories and methylation status. Match columns to the order of columns in the overlap count files
row.names(McapFeatureOverlaps) <- c("Genes", "CDS", "Introns", "Flanking Regions", "Upstream Flanks", "Downstream Flanks", "Intergenic") #Assign row names
head(McapFeatureOverlaps) #Confirm changes
```

```{r}
McapFeatureOverlaps$allCpGs <- c(McapGenomeFeatures$counts[5],
                                 McapGenomeFeatures$counts[1],
                                 McapGenomeFeatures$counts[7],
                                 McapGenomeFeatures$counts[3],
                                 McapGenomeFeatures$counts[4],
                                 McapGenomeFeatures$counts[2],
                                 McapGenomeFeatures$counts[6]) #Assign information for CG motif overlaps with genome features.
head(McapFeatureOverlaps) #Confirm modification
```

```{r}
for (i in 1:length(McapGeneOverlaps$counts)) {
  McapFeatureOverlaps[1,i+1] <- McapGeneOverlaps[i,1]
  McapFeatureOverlaps[2,i+1] <- McapCDSOverlaps[i,1]
  McapFeatureOverlaps[3,i+1] <- McapIntronsOverlaps[i,1]
  McapFeatureOverlaps[4,i+1] <- McapFlanksOverlaps[i,1]
  McapFeatureOverlaps[5,i+1] <- McapFlanksUpstreamOverlaps[i,1]
  McapFeatureOverlaps[6,i+1] <- McapFlanksDownstreamOverlaps[i,1]
  McapFeatureOverlaps[7,i+1] <- McapIntergenicOverlaps[i,1]
} #For each table with feature overlap information, paste the contents of the count column in the assigned row
tail(McapFeatureOverlaps) #Check summary table
```

```{r}
write.table(McapFeatureOverlaps, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap_union-Genomic-Location-Counts.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save file
```

```{r}
McapFeatureOverlapsPercents <- McapFeatureOverlaps[-c(1,4),] #Duplicate dataframe but remove gene and total flank rows
for (i in 1:length(McapFeatureOverlaps)) {
  McapFeatureOverlapsPercents[,i] <- (McapFeatureOverlapsPercents[,i] / (sum(McapFeatureOverlapsPercents[,i]))) * 100
} #Divide every entry by sum of the column and multiply by 100 to get percentages. Do not include gene information
head(McapFeatureOverlapsPercents) #Check calculations
```

```{r}
write.table(McapFeatureOverlapsPercents, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap_union-Genomic-Location-Percents.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save file
```

## Pact

### Import file counts

```{r}
PactGeneOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paGenes-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-gene overlaps
PactGeneOverlaps <- PactGeneOverlaps[-37,] #Remove final row
tail(PactGeneOverlaps) #Confirm import
```

```{r}
PactCDSOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paCDS-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-CDS overlaps
PactCDSOverlaps <- PactCDSOverlaps[-37,] #Remove final row
tail(PactCDSOverlaps) #Confirm import
```

```{r}
PactIntronsOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paIntron-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Introns overlaps
PactIntronsOverlaps <- PactIntronsOverlaps[-37,] #Remove final row
tail(PactIntronsOverlaps) #Confirm import
```

```{r}
PactIntergenicOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paIntergenic-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Intergenic overlaps
PactIntergenicOverlaps <- PactIntergenicOverlaps[-37,] #Remove final row
tail(PactIntergenicOverlaps) #Confirm import
```
