---
title: "Chacterizing CpG Methylation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script, I'll create figures to characterize CpG methylation in my *C. virginica* gonad samples. I will use a list of CpGs with at least 5x coverage in any samples. I'll also calculate relevant statistics for methylation islands.

**MOVE CODE TO ANALYSIS DIRECTORY `2019-03-18-Characterizing-CpG-Methylation` BEFORE PROCEEDING.**

# Session information

```{r}
sessionInfo()
```

# CpG distributions

## Import data

### *M. capitata*

#### Sample 1

```{r}
cpgMethylationWGBS1 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth10.fastp-trim.202003063900_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationWGBS1) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationWGBS1) #Confirm import
```

```{r}
cpgMethylationRRBS1 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth13.fastp-trim.202003062040_R1_001_bismark_bt2_pe_10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationRRBS1) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationRRBS1) #Confirm import
```

```{r}
cpgMethylationMBD1 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth16.fastp-trim.202003062412_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationMBD1) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationMBD1) #Confirm import
```

#### Sample 2

```{r}
cpgMethylationWGBS2 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth11.fastp-trim.202003065734_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationWGBS2) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationWGBS2) #Confirm import
```

```{r}
cpgMethylationRRBS2 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth14.fastp-trim.202003064415_R1_001_bismark_bt2_pe_10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationRRBS2) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationRRBS2) #Confirm import
```

```{r}
cpgMethylationMBD2 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth17.fastp-trim.202003063731_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationMBD2) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationMBD2) #Confirm import
```

#### Sample 3

```{r}
cpgMethylationWGBS3 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth12.fastp-trim.202003060645_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationWGBS3) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationWGBS3) #Confirm import
```

```{r}
cpgMethylationRRBS3 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth15.fastp-trim.202003065503_R1_001_bismark_bt2_pe_10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationRRBS3) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationRRBS3) #Confirm import
```

```{r}
cpgMethylationMBD3 <- read.delim("../analyses/Characterizing-CpG-Methylation/Meth18.fastp-trim.202003065117_R1_001_bismark_bt2_pe._10x.bedgraph-Mcap", header = FALSE) #Import file with CpG methylation for all loci with 10x coverage
colnames(cpgMethylationMBD3) <- c("chromosome", "start", "end", "methylation") #Add column names
head(cpgMethylationMBD3) #Confirm import
```

## Create individual figures to test options

```{r}
hist(cpgMethylationWGBS1$methylation)
hist(cpgMethylationRRBS1$methylation)
hist(cpgMethylationMBD1$methylation)

hist(cpgMethylationWGBS2$methylation)
hist(cpgMethylationRRBS2$methylation)
hist(cpgMethylationMBD2$methylation)

hist(cpgMethylationWGBS3$methylation)
hist(cpgMethylationRRBS3$methylation)
hist(cpgMethylationMBD3$methylation)
```

```{r}
plot(density(cpgMethylationWGBS1$methylation))
plot(density(cpgMethylationRRBS1$methylation))
plot(density(cpgMethylationMBD1$methylation))

plot(density(cpgMethylationWGBS2$methylation))
plot(density(cpgMethylationRRBS2$methylation))
plot(density(cpgMethylationMBD2$methylation))

plot(density(cpgMethylationWGBS3$methylation))
plot(density(cpgMethylationRRBS3$methylation))
plot(density(cpgMethylationMBD3$methylation))
```

## Create multipanel figure with all samples

```{r}
par(mfrow = c(3,3))
hist(cpgMethylationWGBS1$methylation, main = "")
hist(cpgMethylationRRBS1$methylation, main = "")
hist(cpgMethylationMBD1$methylation, main = "")

hist(cpgMethylationWGBS2$methylation, main = "")
hist(cpgMethylationRRBS2$methylation, main = "")
hist(cpgMethylationMBD2$methylation, main = "")

hist(cpgMethylationWGBS3$methylation, main = "")
hist(cpgMethylationRRBS3$methylation, main = "")
hist(cpgMethylationMBD3$methylation, main = "")
```

# Methylation island characteristics

## Import data

```{r}
methylationIslands <- read.delim("../analyses/Characterizing-CpG-Methylation/FILE", header = FALSE) #Import methylation island list
colnames(methylationIslands) <- c("chr", "MI.start", "MI.end", "mCpGs", "MI.length") #Add column names
head(methylationIslands) #Confirm file format
```

## Calculate statistics

```{r}
range(methylationIslands$mCpGs) #Calculate the range of the number of mCpGs
median(methylationIslands$mCpGs) #Calculate the median number of mCpGs in an island
```

```{r}
range(methylationIslands$MI.length) #Calculate the range MI length
median(methylationIslands$MI.length) #Calculate the median MI length
```

