---
title: "Chacterizing CpG Methylation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script, I'll create summary tables and figures to characterize CpG methylation in *M. capitata* and *P. acuta* using WGBS, RRBS, and MBD-BSSeq. I will use a list of CpGs with at least 5x coverage in any samples.

# Session information

```{r}
sessionInfo()
```

# Summary tables

## CpG distribution

### Mcap

#### Import file counts

```{r}
McapAll <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-bedgraph-counts.txt", header = FALSE, col.names = c("totalLines", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapAll <- McapAll[-10,] #Remove last row (total lines for all files)
tail(McapAll) #Confirm import
```

```{r}
McapMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-Meth-counts.txt", header = FALSE, col.names = c("Meth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapMeth <- McapMeth[-10,] #Remove last row (total lines for all files)
tail(McapMeth) #Confirm import
```

```{r}
McapSparseMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-sparseMeth-counts.txt", header = FALSE, col.names = c("sparseMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapSparseMeth <- McapSparseMeth[-10,] #Remove last row (total lines for all files)
tail(McapSparseMeth) #Confirm import
```

```{r}
McapUnMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-unMeth-counts.txt", header = FALSE, col.names = c("unMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
McapUnMeth <- McapUnMeth[-10,] #Remove last row (total lines for all files)
tail(McapUnMeth) #Confirm import
```

#### Create summary table

```{r}
McapCpGType <- cbind(McapAll, McapMeth, McapSparseMeth, McapUnMeth) #Mash tables together by column
McapCpGType <- McapCpGType[,-c(2,4,6,8)] #Remove filename columns
tail(McapCpGType) #Confirm table mashing
```

```{r}
McapCpGType$percentMeth <- (McapCpGType$Meth / McapCpGType$totalLines) * 100 #Calculate percent methylated loci
McapCpGType$percentSparseMeth <- (McapCpGType$sparseMeth / McapCpGType$totalLines) * 100 #Calculate percent sparsely methylated loci
McapCpGType$percentUnMeth <- (McapCpGType$unMeth / McapCpGType$totalLines) * 100 #Calculate percent unmethylated loci
McapCpGType <- McapCpGType[,c(1, 2, 5, 3, 6, 4, 7)] #Reorganize columns
tail(McapCpGType) #Confirm calculations
```

```{r}
write.table(McapCpGType, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CpG-Type.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

### Pact

#### Import file counts

```{r}
PactAll <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-bedgraph-counts.txt", header = FALSE, col.names = c("totalLines", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactAll <- PactAll[-10,] #Remove last row (total lines for all files)
tail(PactAll) #Confirm import
```

```{r}
PactMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-Meth-counts.txt", header = FALSE, col.names = c("Meth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactMeth <- PactMeth[-10,] #Remove last row (total lines for all files)
tail(PactMeth) #Confirm import
```

```{r}
PactSparseMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-sparseMeth-counts.txt", header = FALSE, col.names = c("sparseMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactSparseMeth <- PactSparseMeth[-10,] #Remove last row (total lines for all files)
tail(PactSparseMeth) #Confirm import
```

```{r}
PactUnMeth <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-unMeth-counts.txt", header = FALSE, col.names = c("unMeth", "filename")) #Import file using space as a delimiter. Columns are the number of lines and the filename
PactUnMeth <- PactUnMeth[-10,] #Remove last row (total lines for all files)
tail(PactUnMeth) #Confirm import
```

#### Create summary table

```{r}
PactCpGType <- cbind(PactAll, PactMeth, PactSparseMeth, PactUnMeth) #Mash tables together by column
PactCpGType <- PactCpGType[,-c(2,4,6,8)] #Remove filename columns
tail(PactCpGType) #Confirm table mashing
```

```{r}
PactCpGType$percentMeth <- (PactCpGType$Meth / PactCpGType$totalLines) * 100 #Calculate percent methylated loci
PactCpGType$percentSparseMeth <- (PactCpGType$sparseMeth / PactCpGType$totalLines) * 100 #Calculate percent sparsely methylated loci
PactCpGType$percentUnMeth <- (PactCpGType$unMeth / PactCpGType$totalLines) * 100 #Calculate percent unmethylated loci
PactCpGType <- PactCpGType[,c(1, 2, 5, 3, 6, 4, 7)] #Reorganize columns
tail(PactCpGType) #Confirm calculations
```

```{r}
write.table(PactCpGType, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CpG-Type.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

## CpG overlaps

### Mcap

#### Genes

##### Import file counts

```{r}
McapGeneOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcGenes-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-gene overlaps
McapGeneOverlaps <- McapGeneOverlaps[-37,] #Remove final row
tail(McapGeneOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
McapGeneSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(McapGeneSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(McapGeneSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  McapGeneSummary[i,1] <- McapGeneOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(McapGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapGeneSummary[i,2] <- McapGeneOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(McapGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapGeneSummary[i,3] <- McapGeneOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(McapGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapGeneSummary[i,4] <- McapGeneOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(McapGeneSummary) #Confirm values were added
```

##### Calculate percents

```{r}
McapGeneSummary$percentMeth <- (McapGeneSummary$Meth / McapGeneSummary$totalLines) * 100 #Calculate percent methylated
McapGeneSummary$percentSparseMeth <- (McapGeneSummary$sparseMeth / McapGeneSummary$totalLines) * 100 #Calculate percent sparsely methylated
McapGeneSummary$percentUnMeth <- (McapGeneSummary$unMeth / McapGeneSummary$totalLines) * 100 #Calculate percent unmethylated
McapGeneSummary <- McapGeneSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(McapGeneSummary) #Confirm calculations
```

```{r}
write.table(McapGeneSummary, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-Gene-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### CDS

##### Import file counts

```{r}
McapCDSOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcCDS-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-CDS overlaps
McapCDSOverlaps <- McapCDSOverlaps[-37,] #Remove final row
tail(McapCDSOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
McapCDSSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(McapCDSSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(McapCDSSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  McapCDSSummary[i,1] <- McapCDSOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(McapCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapCDSSummary[i,2] <- McapCDSOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(McapCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapCDSSummary[i,3] <- McapCDSOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(McapCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapCDSSummary[i,4] <- McapCDSOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(McapCDSSummary) #Confirm values were added
```

##### Calculate percents

```{r}
McapCDSSummary$percentMeth <- (McapCDSSummary$Meth / McapCDSSummary$totalLines) * 100 #Calculate percent methylated
McapCDSSummary$percentSparseMeth <- (McapCDSSummary$sparseMeth / McapCDSSummary$totalLines) * 100 #Calculate percent sparsely methylated
McapCDSSummary$percentUnMeth <- (McapCDSSummary$unMeth / McapCDSSummary$totalLines) * 100 #Calculate percent unmethylated
McapCDSSummary <- McapCDSSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(McapCDSSummary) #Confirm calculations
```

```{r}
write.table(McapCDSSummary, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CDS-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### Introns

##### Import file counts

```{r}
McapIntronsOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcIntrons-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Introns overlaps
McapIntronsOverlaps <- McapIntronsOverlaps[-37,] #Remove final row
tail(McapIntronsOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
McapIntronsSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(McapIntronsSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(McapIntronsSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  McapIntronsSummary[i,1] <- McapIntronsOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(McapIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntronsSummary[i,2] <- McapIntronsOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(McapIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntronsSummary[i,3] <- McapIntronsOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(McapIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntronsSummary[i,4] <- McapIntronsOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(McapIntronsSummary) #Confirm values were added
```

##### Calculate percents

```{r}
McapIntronsSummary$percentMeth <- (McapIntronsSummary$Meth / McapIntronsSummary$totalLines) * 100 #Calculate percent methylated
McapIntronsSummary$percentSparseMeth <- (McapIntronsSummary$sparseMeth / McapIntronsSummary$totalLines) * 100 #Calculate percent sparsely methylated
McapIntronsSummary$percentUnMeth <- (McapIntronsSummary$unMeth / McapIntronsSummary$totalLines) * 100 #Calculate percent unmethylated
McapIntronsSummary <- McapIntronsSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(McapIntronsSummary) #Confirm calculations
```

```{r}
write.table(McapIntronsSummary, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-Introns-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### Intergenic regions

##### Import file counts

```{r}
McapIntergenicOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-5x-mcIntergenic-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Intergenic overlaps
McapIntergenicOverlaps <- McapIntergenicOverlaps[-37,] #Remove final row
tail(McapIntergenicOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
McapIntergenicSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(McapIntergenicSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(McapIntergenicSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  McapIntergenicSummary[i,1] <- McapIntergenicOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(McapIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntergenicSummary[i,2] <- McapIntergenicOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(McapIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntergenicSummary[i,3] <- McapIntergenicOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(McapIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  McapIntergenicSummary[i,4] <- McapIntergenicOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(McapIntergenicSummary) #Confirm values were added
```

##### Calculate percents

```{r}
McapIntergenicSummary$percentMeth <- (McapIntergenicSummary$Meth / McapIntergenicSummary$totalLines) * 100 #Calculate percent methylated
McapIntergenicSummary$percentSparseMeth <- (McapIntergenicSummary$sparseMeth / McapIntergenicSummary$totalLines) * 100 #Calculate percent sparsely methylated
McapIntergenicSummary$percentUnMeth <- (McapIntergenicSummary$unMeth / McapIntergenicSummary$totalLines) * 100 #Calculate percent unmethylated
McapIntergenicSummary <- McapIntergenicSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(McapIntergenicSummary) #Confirm calculations
```

```{r}
write.table(McapIntergenicSummary, "../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-Intergenic-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

### Pact

#### Genes

##### Import file counts

```{r}
PactGeneOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paGenes-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-gene overlaps
PactGeneOverlaps <- PactGeneOverlaps[-37,] #Remove final row
tail(PactGeneOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
PactGeneSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(PactGeneSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(PactGeneSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  PactGeneSummary[i,1] <- PactGeneOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(PactGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactGeneSummary[i,2] <- PactGeneOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(PactGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactGeneSummary[i,3] <- PactGeneOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(PactGeneSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactGeneSummary[i,4] <- PactGeneOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(PactGeneSummary) #Confirm values were added
```

##### Calculate percents

```{r}
PactGeneSummary$percentMeth <- (PactGeneSummary$Meth / PactGeneSummary$totalLines) * 100 #Calculate percent methylated
PactGeneSummary$percentSparseMeth <- (PactGeneSummary$sparseMeth / PactGeneSummary$totalLines) * 100 #Calculate percent sparsely methylated
PactGeneSummary$percentUnMeth <- (PactGeneSummary$unMeth / PactGeneSummary$totalLines) * 100 #Calculate percent unmethylated
PactGeneSummary <- PactGeneSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(PactGeneSummary) #Confirm calculations
```

```{r}
write.table(PactGeneSummary, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-Gene-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### CDS

##### Import file counts

```{r}
PactCDSOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paCDS-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-CDS overlaps
PactCDSOverlaps <- PactCDSOverlaps[-37,] #Remove final row
tail(PactCDSOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
PactCDSSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(PactCDSSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(PactCDSSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  PactCDSSummary[i,1] <- PactCDSOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(PactCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactCDSSummary[i,2] <- PactCDSOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(PactCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactCDSSummary[i,3] <- PactCDSOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(PactCDSSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactCDSSummary[i,4] <- PactCDSOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(PactCDSSummary) #Confirm values were added
```

##### Calculate percents

```{r}
PactCDSSummary$percentMeth <- (PactCDSSummary$Meth / PactCDSSummary$totalLines) * 100 #Calculate percent methylated
PactCDSSummary$percentSparseMeth <- (PactCDSSummary$sparseMeth / PactCDSSummary$totalLines) * 100 #Calculate percent sparsely methylated
PactCDSSummary$percentUnMeth <- (PactCDSSummary$unMeth / PactCDSSummary$totalLines) * 100 #Calculate percent unmethylated
PactCDSSummary <- PactCDSSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(PactCDSSummary) #Confirm calculations
```

```{r}
write.table(PactCDSSummary, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CDS-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### Introns

##### Import file counts

```{r}
PactIntronsOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paIntron-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Introns overlaps
PactIntronsOverlaps <- PactIntronsOverlaps[-37,] #Remove final row
tail(PactIntronsOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
PactIntronsSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(PactIntronsSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(PactIntronsSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  PactIntronsSummary[i,1] <- PactIntronsOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(PactIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntronsSummary[i,2] <- PactIntronsOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(PactIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntronsSummary[i,3] <- PactIntronsOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(PactIntronsSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntronsSummary[i,4] <- PactIntronsOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(PactIntronsSummary) #Confirm values were added
```

##### Calculate percents

```{r}
PactIntronsSummary$percentMeth <- (PactIntronsSummary$Meth / PactIntronsSummary$totalLines) * 100 #Calculate percent methylated
PactIntronsSummary$percentSparseMeth <- (PactIntronsSummary$sparseMeth / PactIntronsSummary$totalLines) * 100 #Calculate percent sparsely methylated
PactIntronsSummary$percentUnMeth <- (PactIntronsSummary$unMeth / PactIntronsSummary$totalLines) * 100 #Calculate percent unmethylated
PactIntronsSummary <- PactIntronsSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(PactIntronsSummary) #Confirm calculations
```

```{r}
write.table(PactIntronsSummary, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-Introns-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

#### Intergenic regions

##### Import file counts

```{r}
PactIntergenicOverlaps <- read.table("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-5x-paIntergenic-counts.txt", header = FALSE, col.names = c("counts", "filename")) #Import file with all file-Intergenic overlaps
PactIntergenicOverlaps <- PactIntergenicOverlaps[-37,] #Remove final row
tail(PactIntergenicOverlaps) #Confirm import
```

##### Transfer values to summary table

```{r}
PactIntergenicSummary <- data.frame("totalLines" = rep(0, times = 9),
                              "Meth" = rep(0, times = 9),
                              "sparseMeth" = rep(0, times = 9),
                              "unMeth" = rep(0, times = 9)) #Create empty dataframe for final summary table
head(PactIntergenicSummary) #Confirm dataframe creation
```

```{r}
nLines <- length(PactIntergenicSummary$totalLines) #Count number of lines
for(i in 1:nLines) {
  PactIntergenicSummary[i,1] <- PactIntergenicOverlaps[i+(3*i),1]
} #For every file with the total number of CpG with data (i + (3*i)), take the value in the first column (total lines) and save it in the first column
head(PactIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntergenicSummary[i,2] <- PactIntergenicOverlaps[(i+(3*i)-3),1]
} #For every file with the total number of CpG with data ((i + (3*i))-3), take the value in the first column (total lines) and save it in the second column
head(PactIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntergenicSummary[i,3] <- PactIntergenicOverlaps[(i+(3*i)-2),1]
} #For every file with the total number of CpG with data ((i + (3*i))-2), take the value in the first column (total lines) and save it in the third column
head(PactIntergenicSummary) #Confirm values were added
```

```{r}
for(i in 1:nLines) {
  PactIntergenicSummary[i,4] <- PactIntergenicOverlaps[(i+(3*i)-1),1]
} #For every file with the total number of CpG with data ((i + (3*i))-1), take the value in the first column (total lines) and save it in the fourth column
head(PactIntergenicSummary) #Confirm values were added
```

##### Calculate percents

```{r}
PactIntergenicSummary$percentMeth <- (PactIntergenicSummary$Meth / PactIntergenicSummary$totalLines) * 100 #Calculate percent methylated
PactIntergenicSummary$percentSparseMeth <- (PactIntergenicSummary$sparseMeth / PactIntergenicSummary$totalLines) * 100 #Calculate percent sparsely methylated
PactIntergenicSummary$percentUnMeth <- (PactIntergenicSummary$unMeth / PactIntergenicSummary$totalLines) * 100 #Calculate percent unmethylated
PactIntergenicSummary <- PactIntergenicSummary[,c(1,2,5,3,6,4,7)] #Rearrange columns
tail(PactIntergenicSummary) #Confirm calculations
```

```{r}
write.table(PactIntergenicSummary, "../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-Intergenic-Overlaps.txt", sep = "\t", quote = FALSE, row.names = TRUE) #Save summary table
```

# CpG distributions

## Individual samples

### Mcap

```{r}
McapCpGType <- read.delim("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CpG-Type.txt", sep = "\t", header = TRUE) #Import summary table
head(McapCpGType) #Check import
```

#### Reorganize data

```{r}
McapCpGTypePercents <- McapCpGType[,c(3,5,7)] #Keep only columns with % total CpG information
head(McapCpGTypePercents) #Check reorganization
```

```{r}
#Create test plots

barplot(t(McapCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(McapCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, RRBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(McapCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, MBD-BSSeq
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")
```

#### Create multipanel figure with all samples

```{r}
#pdf("../analyses/Characterizing-CpG-Methylation-5x/Mcap/Mcap-CpG-Type.pdf", height = 8.5, width = 11) #Save file as pdf

par(mfcol = c(3,3), mar = c(2, 2, 2, 0), oma = c(5, 5, 2, 0)) #Fill in multipanel plot by column and adjust inner and outer margins

barplot(t(McapCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 3, "WBGS", adj = 0, line = 1) #Add sequencing information

barplot(t(McapCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, RRBS
mtext(side = 3, "RRBS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 2, outer = TRUE, "% 5x CpG with Data", cex = 1.5, line = 2) #Add y-axis label

barplot(t(McapCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 1, MBD-BSSeq
mtext(side = 3, "MBD-BSSeq", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(McapCpGTypePercents[2,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, WGBS
barplot(t(McapCpGTypePercents[5,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, RRBS
barplot(t(McapCpGTypePercents[8,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 2, MBD-BSSeq
mtext(side = 1, outer = TRUE, "% Methylation", cex = 1.5, line = 2) #Add x-axis label

barplot(t(McapCpGTypePercents[3,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, WGBS
barplot(t(McapCpGTypePercents[6,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, RRBS
barplot(t(McapCpGTypePercents[9,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 3, MBD-BSSeq

#dev.off() #Turn off plotting device
```

### Pact

#### Reorganize data

```{r}
PactCpGTypePercents <- PactCpGType[,c(3,5,7)] #Keep only columns with % total CpG information
tail(PactCpGTypePercents) #Check reorganization
```

```{r}
#Create test plots

barplot(t(PactCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, WGBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(PactCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, RRBS
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")

barplot(t(PactCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c("Methylated", "Sparsely Methylated", "Unmethylated"), axes = FALSE) #Sample 1, MBD-BSSeq
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80")
```

#### Create multipanel figure with all samples

```{r}
#pdf("../analyses/Characterizing-CpG-Methylation-5x/Pact/Pact-CpG-Type.pdf", height = 8.5, width = 11) #Save file as pdf

par(mfcol = c(3,3), mar = c(2, 2, 2, 0), oma = c(5, 5, 2, 0)) #Fill in multipanel plot by column and adjust inner and outer margins

barplot(t(PactCpGTypePercents[1,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, WGBS
mtext(side = 3, "WBGS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(PactCpGTypePercents[4,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 1, RRBS
mtext(side = 3, "RRBS", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis
mtext(side = 2, outer = TRUE, "% 5x CpG with Data", cex = 1.5, line = 2) #Add y-axis label

barplot(t(PactCpGTypePercents[7,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 1, MBD-BSSeq
mtext(side = 3, "MBD-BSSeq", adj = 0, line = 1) #Add sequencing information
axis(side = 2, at = seq(0, 100, by = 25), las = 2, col = "grey80", cex.axis = 1.3) #Add y-axis

barplot(t(PactCpGTypePercents[2,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, WGBS
barplot(t(PactCpGTypePercents[5,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 2, RRBS
barplot(t(PactCpGTypePercents[8,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 2, MBD-BSSeq
mtext(side = 1, outer = TRUE, "% Methylation", cex = 1.5, line = 2) #Add x-axis label

barplot(t(PactCpGTypePercents[3,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, WGBS
barplot(t(PactCpGTypePercents[6,]), beside = TRUE, ylim = c(0,100), names.arg = FALSE, axes = FALSE) #Sample 3, RRBS
barplot(t(PactCpGTypePercents[9,]), beside = TRUE, ylim = c(0,100), names.arg = c(expression("">= "50%"), "10-50%", "< 10%"), cex.names = 1.3, axes = FALSE) #Sample 3, MBD-BSSeq

#dev.off() #Turn off plotting device
```

## Combined samples

### Mcap

#### Reorganize data

```{r}
McapCpGTypeCounts <- data.frame("totalLines" = rep(0, times = 3),
                                   "Meth" = rep(0, times = 3),
                                   "sparseMeth" = rep(0, times = 3),
                                   "unMeth" = rep(0, times = 3)) #Create an empty dataframe for summed counts
head(McapCpGTypeCounts) #Check dataframe format
```

```{r}
#Total lines concatenation
McapCpGTypeCounts$totalLines[1] <- sum(McapCpGType$totalLines[1], McapCpGType$totalLines[2],  McapCpGType$totalLines[3]) #WGBS
McapCpGTypeCounts$totalLines[2] <- sum(McapCpGType$totalLines[4], McapCpGType$totalLines[5],  McapCpGType$totalLines[6]) #RRBS
McapCpGTypeCounts$totalLines[3] <- sum(McapCpGType$totalLines[7], McapCpGType$totalLines[8],  McapCpGType$totalLines[9]) #MBD-BSSeq
head(McapCpGTypeCounts) #Check output
```

```{r}
#Meth concatenation
McapCpGTypeCounts$Meth[1] <- sum(McapCpGType$Meth[1], McapCpGType$Meth[2],  McapCpGType$Meth[3]) #WGBS
McapCpGTypeCounts$Meth[2] <- sum(McapCpGType$Meth[4], McapCpGType$Meth[5],  McapCpGType$Meth[6]) #RRBS
McapCpGTypeCounts$Meth[3] <- sum(McapCpGType$Meth[7], McapCpGType$Meth[8],  McapCpGType$Meth[9]) #MBD-BSSeq
head(McapCpGTypeCounts) #Check output
```

```{r}
#sparseMeth concatenation
McapCpGTypeCounts$sparseMeth[1] <- sum(McapCpGType$sparseMeth[1], McapCpGType$sparseMeth[2],  McapCpGType$sparseMeth[3]) #WGBS
McapCpGTypeCounts$sparseMeth[2] <- sum(McapCpGType$sparseMeth[4], McapCpGType$sparseMeth[5],  McapCpGType$sparseMeth[6]) #RRBS
McapCpGTypeCounts$sparseMeth[3] <- sum(McapCpGType$sparseMeth[7], McapCpGType$sparseMeth[8],  McapCpGType$sparseMeth[9]) #MBD-BSSeq
head(McapCpGTypeCounts) #Check output
```

```{r}
#unMeth concatenation
McapCpGTypeCounts$unMeth[1] <- sum(McapCpGType$unMeth[1], McapCpGType$unMeth[2],  McapCpGType$unMeth[3]) #WGBS
McapCpGTypeCounts$unMeth[2] <- sum(McapCpGType$unMeth[4], McapCpGType$unMeth[5],  McapCpGType$unMeth[6]) #RRBS
McapCpGTypeCounts$unMeth[3] <- sum(McapCpGType$unMeth[7], McapCpGType$unMeth[8],  McapCpGType$unMeth[9]) #MBD-BSSeq
head(McapCpGTypeCounts) #Check output
```

```{r}
McapCpGTypeTotalPercents <- data.frame("Meth" = rep(0, times = 3),
                                       "sparseMeth" = rep(0, times = 3),
                                       "unMeth" = rep(0, times = 3)) #Create an empty dataframe for total percents
head(McapCpGTypeTotalPercents) #Check dataframe
```

```{r}
for (i in 1:3) {
  McapCpGTypeTotalPercents[,i] <- (McapCpGTypeCounts[,i+1] / McapCpGTypeCounts[,1])  *100
} #Divide CpG type by total lines and multiply by 100 to get percents
head(McapCpGTypeTotalPercents) #Check output
```

#### Create plot

```{r}
RColorBrewer::display.brewer.all()
```


```{r}
barplot(t(McapCpGTypeTotalPercents))
```


### Pact














